---
title: "Cleaning procedure"
author: "Paul Rougieux"
date: "04/12/2014"
output: 
  html_document:
    toc: yes
---
## Purpose
For wood products, each trade flows 
should have an estimated volume in m3
For paper products, each trade flwos
should have an estimated weight in kg

Each tradeflow goes through 2 filters: 
1. Handling missing data
    + If the weight is available in kg. Estimate missing quantity from weight using a regional conversion factor. 
    + If neither volume nor quantity is available, use a regional unit price to convert trade value to a quantity.
2. Handling out of bounds data
    + Check if the unit price is above lowerprice and below upperprice
    correct using medianprice if out of bounds


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
library(tradeflows)
library(dplyr)
library(RMySQL)
``` 


## Calculate without pre-made functions
Load data from a file. (in the final system data will be loaded from a database)
Calculate median unit price by region and by year.
Calculate median conversion factor by year for the world.
Median price calculation can be put in a function for further use.
```{r loadrawdata}
load("data-raw/comtrade/440799.RData")
dtfraw <- dtf %>% renamecolumns %>% removeduplicatedflows

# Quantity not available but weight is available
dtfraw %>% filter(is.na(quantity) & !is.na(weight)) %>% nrow
# Quantity and weight not available
dtfraw %>% filter(is.na(quantity) & is.na(weight)) %>% nrow

dtf <- dtfraw %>% 
    filter(flow %in% c("Import", "Export")) %>%
    # Remove EU28 reporter
    filter(!reporter %in%c("EU-28")) %>%
    # Remove EU-28 and World partner
    filter(!partner %in%c("EU-28", "World")) %>%
    # Add regionreporter and regionpartner
    merge(select(reportercomtrade, reportercode, regionreporter=region), 
          all.x=TRUE) %>%
    merge(select(reportercomtrade, 
                 partnercode = reportercode, regionpartner=region),
          all.x=TRUE) %>%
    mutate(conversion = weight / quantity,
           price = tradevalue / quantity,
           # To avoid "integer overflow - use sum(as.numeric(.))" error 
           # on sum of all values
           tradevalue = as.numeric(tradevalue),
           # To avoid Error in dtf$c(NA_integer_,  : 
           # invalid subscript type 'integer'
           quantity = as.numeric(quantity))
```


Check for unique flows
```{r}
nrow(dtf)
dtf %>% select(year, flow, reportercode, partnercode, 
               productcode, classification) %>% unique %>%nrow
message("There were duplicated lines for the following reporters:")
dtf %>% findduplicatedflows %>% select(year, reporter) %>% unique %>%
    arrange(reporter) %>% kable
```

### Prepare Unit prices and conversion factors

#### Include (or not) comtrade estimates of quantity from the unit price calculation
Check filter on flag below to see if comtrade estimates are included or not.
```{r extractpricesandconversionfactors}


pricesregion <- dtf %>%
    filter(!is.na(quantity) & unit !="No Quantity") %>%
    # Exclude comtrade quantity estimates
#     filter(flag==0 |flag==4 ) %>%
    group_by(flow, regionreporter, year, unit) %>%
    summarise(lowerprice = round(0.5 * quantile(price, 0.25, 
                                                names=FALSE, na.rm=TRUE)),
              medianprice = round(median(price, na.rm=TRUE)),
              upperprice = round(2 * quantile(price, 0.75,
                                              names=FALSE, na.rm=TRUE))) %>%
    arrange(-medianprice)

conversionfactorworld <- dtf %>%
    filter(!is.na(quantity) & unit !="No Quantity") %>%
    group_by(flow, year, unit) %>%
    summarise(medianconversion = round(median(conversion,na.rm=TRUE)))
```

### Handling missing quantity data
* If the quantity is in kg or m3 take the quantity as is
* If the weight in kilo is available, 
   convert using one conversion factor for the entire world.
* If neither volume nor quantity is available, 
   use a regional unit price to calculate the quantity
```{r handlingmissingquantity}
# Quantity based on the weight using a conversion factor
dtf <- merge(dtf, conversionfactorworld, all.x=TRUE) %>%
    mutate(quantity_cf = weight / medianconversion)
# Quantity based on the trade value, using a unit price
dtf <- merge(dtf, pricesregion, all.x=TRUE) %>%
    mutate(quantity_up = tradevalue / medianprice,
           havequantity = !is.na(quantity)) 

summarychange  <- function(dtf){
    dtf %>% summarise(rows = n(),
                      quantity = sum(quantity, na.rm=TRUE) ,
                      quantity_cf = sum(quantity_cf, na.rm=TRUE),
                      cf_increase = round((quantity_cf-quantity)/ quantity *100),
                      quantity_up = sum(quantity_up, na.rm=TRUE),
                      up_increase = round((quantity_up-quantity)/ quantity *100))
    }

dtf %>% group_by(havequantity, year, flow) %>% summarychange %>% kable
dtf %>% group_by(havequantity, flag, flow) %>% summarychange %>% kable
dtf %>% group_by(flag) %>% summarychange %>% kable
dtf %>% group_by(round(log(quantity)), flag) %>% summarychange %>% kable

# Check if both tables contain the same total world trade flows


print("Percent change in world trade flows ")
wtf <- sum(dtf$quantity,na.rm=TRUE)
dtf %>% 
    #filter(!havequantity) %>% 
    group_by(havequantity, year, flow, flag) %>%
    summarise(changeinworldtrade_cf = round((sum(quantity_cf,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / wtf*100,3),
              changeinworldtrade_up = 
                  round((sum(quantity_up,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / 
                            wtf*100,3))  %>%
    kable
```

### Handling price bounds
```{r handlingpricebounds}

```


### Handling partner information
```{r handlingpartner}

```


## Calculate with functions from clean.R
### Compare unit prices and conversion factor tables
Depending on whether quantity estimates have been removed or not, 
unit prices will be very different.
```{r cleanRextractprices}
dtf2 <- dtfraw %>% 
    addconversionfactorandprice %>%
    addregion
#     merge(select(reportercomtrade, 
#                  partnercode = reportercode, regionpartner=region))    
# Check for duplicated flows
nrow(findduplicatedflows(dtf2))
# Extract prices and conversion factors
pricesregion2 <- dtf2 %>% extractprices(includeqestimates = TRUE)
conversionfactorworld2 <- dtf2 %>% 
    extractconversionfactors(includeqwestimates = TRUE)

# Compare with other chunk
summary(pricesregion$medianprice - pricesregion2$medianprice)
summary(conversionfactorworld$medianconversion -
            conversionfactorworld2$medianconversion)
```

### Handling missing quantity data and price bounds
```{r cleanRhandlingmissingquantity}
dtfq <- dtfraw %>% 
    addconversionfactorandprice %>% 
    addregion %>%
    estimatequantity(price=pricesregion2, 
                     conversionfactor=conversionfactorworld2, shaveprice=FALSE)
nrow(findduplicatedflows(dtfq))

fr2 <- dtfq %>% filter(reporter =="France") #%>% removeduplicatedflows
# Check flags
fr <- estimatequantity(dtf, price=pricesregion2, 
                       conversionfactor=conversionfactorworld2) %>%
    filter(reporter=="France") 
fr2$partner[!fr2$partner %in% fr$partner]
# World makes the difference
dtfqn <- dtfq %>% group_by(reporter) %>% summarise(n=n()) %>% arrange(-n)
dtfn <- dtf %>% group_by(reporter) %>% summarise(n=n())%>% arrange(-n)
fr1 <- dtf %>% filter(reporter=="France")
```

### Handling partner information
```{r cleanRhandlingpartner}
dtf_p <- addpartnerflow(dtf)
```


### Main clean function
```{r mainclean}
dtfclean <- dtfraw %>% 
    clean(shaveprice=TRUE) 
dtfdup <- dtfclean %>% findduplicatedflows
dtfclean %>% addpartnerflow
dtfclean %>% group_by(flag) %>% summarychange %>% kable
```

## Clean from the database
### Read from the database, clean, write to the database
```{r readfromdb, eval=FALSE}
charcoal90 <- loadrawdata(440290)
# Add missing fields as of January 2015 # should be fixed later
charcoal90$flag <- 0
# Check that all required columns are there
stopifnot(column_names$efi[column_names$raw_flow] %in% names(dtf))
missingcolumnsraw <- column_names$efi[column_names$raw_flow &
                                          !column_names$efi %in% names(dtf)]
if (length(missingcolumnsraw)>0){
    warning("Missing columns in the raw_flow table: \n",
            paste(missingcolumnsraw, collapse=", "))
}
charcoal90 <- clean(charcoal90, deleteextracolumns = FALSE)
```


```{r readfromdb2, eval=FALSE}
```


### Clean in detail again using all functions 
```{r cleandbfunction, eval=FALSE}
swd99 <- readdbproduct(440799, dbtable = "raw_flow_yearly") %>% 
    removeduplicatedflows %>%
    addconversionfactorandprice %>%
    addregion
priceregion <- extractprices(swd99)
conversionfactorworld <- extractconversionfactors(swd99)
swd99 <- swd99 %>%
        estimatequantity(priceregion, conversionfactorworld,
                         shaveprice = TRUE) %>%
        addpartnerflow
names(swd99)
deletedbproduct(440799, dbtable= "validated_flow_yearly")
writedbproduct(swd99, dbtable = "validated_flow_yearly" )
```


### Clean using the main clean function
```{r cleanfromdbmain, eval = FALSE}
swd99 <- readdbproduct(440799, dbtable = "raw_flow_yearly")
rawnames <- names(swd99)
swd99 <- clean(swd99)
# Only 2 new columns because clean removed extracolumns
names(swd99)[!names(swd99) %in% rawnames]
writedbproduct(swd99, dbtable = "validated_flow_yearly" )
```


### Cleandb 
```{r cleandb}
cleandbproduct(440799, tableread =  "raw_flow_yearly", tablewrite = "validated_flow_yearly")

# Wrong writetable to test failure
cleandbproduct(440799, tableread =  "raw_flow_yearly", tablewrite = "validated_flow_yearly_wrong")

```


### Other experiments with dbwrite
```{r writetodbexperimental, eval=FALSE}
setdatabaseconfig(message=FALSE)
require(RMySQL)
db <- getOption("tradeflowsDB")
DBwrite <- dbConnect(MySQL(), user=db["user"], host=db["host"],
                     password=db["password"], dbname=db["dbname"])
result <- dbWriteTable(DBwrite, name="validated_flow_yearly",
                       value=data.frame(swd99), append=TRUE, row.names = FALSE)
dbDisconnect(DBwrite)

writedbpaul <-function(dtf, dbtable){
    # Write only to a validated_flow table
    stopifnot(dbtable %in% c("validated_flow_yearly", "validated_flow_monthly"))
    setdatabaseconfig(message=FALSE)
    require(RMySQL)
    db <- getOption("tradeflowsDB")
    DBwrite <- dbConnect(MySQL(), user=db["user"], host=db["host"],
                         password=db["password"], dbname=db["dbname"])
    result <- dbWriteTable(DBwrite, name="validated_flow_yearly",
                           value=data.frame(dtf), append=TRUE, row.names = FALSE)
    dbDisconnect(DBwrite)
    return(result)
    }
writedbpaul(swd99, "validated_flow_yearly")


writedbproduct <- function(dtf, dbtable){
    # Write only to a validated_flow table
    stopifnot(dbtable %in% c("validated_flow_yearly", "validated_flow_monthly"))
    setdatabaseconfig(message=FALSE)
    require(RMySQL)
    db <- getOption("tradeflowsDB")
    DBwrite <- dbConnect(MySQL(), user=db["user"], host=db["host"],
                         password=db["password"], dbname=db["dbname"])
    dtf <- data.frame(dtf)
    result <- dbWriteTable(DBwrite, name=table,
                           value=dtf, append=TRUE, row.names = FALSE)
    dbDisconnect(DBwrite)
    return(result)
}
writedbproduct(swd99, "validated_flow_yearly")
```


## Misceleanous remarks
### Issue with duplicated Sudan in reportercomtrade
```{r duplicatedsudaninreportercomtrade, eval=FALSE}
message("There were duplicated lines for the following reporters:")
# For example this flow reported by Sudan is duplicated
dtf %>% filter(reporter=="Fmr Sudan" & partner=="USA") %>% select(-productdescription)

tfraw %>% renamecolumns %>%
      merge(select(reportercomtrade, 
                 partnercode = reportercode, regionpartner=region)) %>%
    duplicates %>% select(year, reporter, partner) %>% unique %>%
    arrange(reporter) %>% kable

# What happens when we 
#       merge(select(reportercomtrade, 
#                  partnercode = reportercode, regionpartner=region)) 
bli <- dtfraw %>% addconversionfactorandprice %>%
      merge(select(reportercomtrade, 
                 partnercode = reportercode, regionpartner=region)) %>%
    duplicates
# For example this flow reported by France is duplicated
dtf %>% filter(reporter=="France" & partner=="Fmr Sudan") %>% select(-productdescription)
#  partnercode reportercode classification year flowcode   flow reporter reporteriso
# 1         736          251             H2 2005        2 Export   France         FRA
# 2         736          251             H2 2005        2 Export   France         FRA
#     partner partneriso productcode unitcode                   unit quantity weight
# 1 Fmr Sudan        SDN      440799       12 Volume in cubic meters        2   2000
# 2 Fmr Sudan        SDN      440799       12 Volume in cubic meters        2   2000
#   tradevalue flag regionreporter regionpartner conversion price
# 1       2488    0         Europe        Africa       1000  1244
# 2       2488    0         Europe        Africa       1000  1244

# Issue with duplicated Sudan reportercode 729 and Fmr Sudan reportercode 729
# in the reportercomtrade table with causes duplicates in the merged tables
reportercomtrade %>% filter(reportercode %in% c(736,729))

# Are there other duplicates in the reportercomtrade table?
reportercomtrade %>% filter(duplicated(reportercode))  
#  reportercode  reporter reporteriso reportercodefao regionfao       subregion
# 1          729     Sudan         SDN             206    Africa Northern Africa
# 2          736 Fmr Sudan         SDN             206    Africa Northern Africa
#   reporternamefao region duplicate
# 1       the Sudan Africa      TRUE
# 2       the Sudan Africa      TRUE

# This is due to a merge based on duplicated iso codes for Sudan
filter(reportercomtrade, !is.na(reporteriso) & duplicated(reporteriso))
#   reportercode  reporter reporteriso reportercodefao regionfao       subregion
# 1          729     Sudan         SDN             206    Africa Northern Africa
# 2          736 Fmr Sudan         SDN             276    Africa Northern Africa
# 3          736 Fmr Sudan         SDN             206    Africa Northern Africa
#   reporternamefao region
# 1       the Sudan Africa
# 2       the Sudan Africa
# 3       the Sudan Africa

```

```{r beforeremovingduplicates, eval=FALSE}
nrow(dtf)
# [1] 47493
nrow(dtf2)
# [1] 49082
# After removing duplicates
 nrow(dtf)
# [1] 47351
nrow(dtf2)
# [1] 48940
``` 

### Units
In 2004 volume was reporter in litres
```{r}
dtf %>% filter( unit=="Volume in litres") %>% select(year) %>% unique
```

### Unit No quantity creates problem while merging price and conversion factor tables

```{r issueflag200, eval=FALSE}
# look at an issue with the absence of median price and conversion factor in 
# flag 200
dtf200 <- dtfclean %>% filter(flag==200)
dtf200 <- dtf200 %>% merge(conversionfactorworld, all.x=TRUE)
# That was meainly because of this unit "No Quantity"
```
