---
title: "Ratio Weight / Volume and use as a conversion factors"
author: "Paul Rougieux"
date: "04/12/2014"
output: html_document
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 


To calculate conversion factors for sawnwood products,
we divided the weight by the trade volume. 
Then we analysed median conversion factors at the regional and subregional level.
Mean values could not be used since there are a number of infinite 
ratios (due to flows for which the reported volume was 0).

```{r loadandprepare}
load("data-raw/comtrade/440799.RData")
swd99 <- dtf %>% 
    renamecolumns %>% changecolumntype %>% 
    filter(flow %in% c("Import", "Export")) %>%
    # Remove EU28 reporter
    filter(!reporter %in%c("EU-28")) %>%
    # Remove World partner
    filter(!partner %in%c("EU-28", "World")) %>%
    merge(select(reportercomtrade, reportercode, region, subregion)) %>%
    mutate(conversion = weight / quantity,
           price = tradevalue / quantity)
```


### World median conversion factor (kg/m3)
World median conversion factor (kg/m3) by year.
```{r worldmedianbyyear}
swd99 %>% group_by(year, flow) %>%
    summarise(medianconv = round(median(conversion, na.rm=TRUE))) %>%
    dcast(year~flow, value.var="medianconv") %>% kable
```


World median conversion factor (kg/m3) by year, with flag.
```{r worldmedianbyyearwithflag}
swd99 %>% group_by(year, flow, flag) %>%
    summarise(medianconv = round(median(conversion, na.rm=TRUE))) %>%
    dcast(year~flag+flow, value.var="medianconv") %>% kable
```


Netherlands and Belgium
```{r}
swd99 %>% filter(reporter %in% c("Netherlands", "Belgium")) %>%
    group_by(reporter, year, flow, flag) %>%
    summarise(medianconv = round(median(conversion, na.rm=TRUE))) %>%
    dcast(reporter+year~flag+flow, value.var="medianconv") %>% kable
```


### Comtrade estimates and missing quantity data
quantity estimation only 
4 = net weight estimation only 
6 = both quantity and net weight are estimated.
```{r}
flagname <- c("No estimate",
              "Quantity estimate",
              "Weight estimate",
              "Both quandity and weight estimate")
flag <- data.frame(flag=c(0,2,4,6),
                   flagname = ordered(flagname, levels=flagname))
swd99 %>% merge(flag) %>%
    ggplot(aes(x=flagname,y=quantity)) + facet_grid(flow~.) +
    geom_bar(stat="identity") + ggtitle("Estimates in terms of Quantity")

swd99 %>% merge(flag) %>%
    ggplot(aes(x=flagname,y=weight)) + facet_grid(flow~.) +
    geom_bar(stat="identity")  + ggtitle("Estimates in terms of Weight")

swd99 %>% merge(flag) %>%
    ggplot(aes(x=flagname,y=tradevalue)) + facet_grid(flow~.) +
    geom_bar(stat="identity")  + ggtitle("Estimates in terms of Trade value")


swd99 %>% 
    ggplot(aes(x=flag, y=tradevalue, fill=region)) +
    geom_bar(stat="identity") + ggtitle("Estimates in terms of trade value") + 
    facet_grid(region+flow~year)
```



### Distribution of conversion factors between `r min(swd99$year)` and `r max(swd99$year)`
```{r}
p <- ggplot(swd99, aes(x=conversion, fill=region)) + 
    geom_histogram(binwidth=100) + xlim(0,2500) +
    ylab("Number of trade flows") + 
    xlab("Weight / Volume ratio (in kg/m3)")
# p + facet_grid(~flow) + ggtitle("All world flows for 440799")
p + facet_grid(region~flow) + ggtitle("World flows for 440799 by regions")
# p + facet_grid(year~flow) + ggtitle("World flows for 440799 by year")

p <- ggplot(swd99, aes(x=conversion)) + 
    geom_histogram(binwidth=50) +  xlim(0,1500) +
    geom_vline(aes(xintercept=700),color="green") +
    ggtitle("Distribution of conversion factor, number of flows by region. 
            Green line represents conversion factor for
            non coniferous sawnwood 700kg/m3")
p + facet_grid(flow~.) 
p + facet_grid(region+flow~year) 

```


### Conversion factors by year and by trade weight
Conversion factors above 2000 are clear outliers, 
x axis was limited to 2000. 
`r round(nrow(filter(swd99,conversion>2000))/nrow(swd99)*100)`% of the flows
(percentage of the number of flows) had a conversion factor above this limit.
And amoung them,
`r round(nrow(filter(swd99,conversion==Inf))/nrow(swd99)*100)`%
had a conversion factor equal to the infinite value.
```{r conversionweight}
nrow(filter(swd99,conversion==Inf))
# Variable used for facetting
swd99$logweight <- round(log(swd99$weight))
p <- ggplot(filter(swd99,conversion!=Inf), 
       aes(x = conversion, y = weight, color=as.factor(flag))) +
    geom_point(alpha = 1/5) +
    coord_cartesian(xlim =c(0,2000)) +
    ggtitle("Variation of the conversion factor along the weight")
p  
p + scale_y_log10() + ylab("Weight (log scale)")
p + scale_y_log10() + facet_wrap(~year) + ylab("Weight (log scale)")
p + coord_cartesian(xlim =c(0,2000), ylim = c(0,10000)) 
p + facet_wrap(~logweight, scales="free_y") + ylab("Weight split by log(weight)")
p + scale_y_log10() + facet_grid(flag~year) 

# Investigate those straight lines
swd99 %>% filter(logweight == 7 & conversion>1000 & conversion<2000) %>%
    ggplot(aes(x = conversion, y = weight, 
               label=reporter, color=as.factor(flag))) + 
    geom_text() + facet_grid(year ~ flow )
# When both quantity and weight were corrected
swd99 %>% filter(flag == 6) %>%
    ggplot(aes(x=conversion, y = weight)) +
    geom_point(alpha=1/5) + coord_cartesian(xlim=c(0,2000)) +
    scale_y_log10() + facet_grid(region~year)
# Europe 2009
swd99 %>% filter(flag == 6 & region=="Europe" & year==2009) %>%
    ggplot(aes(x=conversion, y = weight, label=reporter)) +
    geom_text() + scale_y_log10()
# Netherlands and Belgium
swd99 %>% filter(flag == 6 & reporter %in% c("Netherlands", "Belgium")) %>%
    ggplot(aes(x=conversion, y = weight, label=partner, color=reporter)) +
    geom_text() + scale_y_log10() + facet_wrap(~year)


# aggregate by country and flow
# swd99agg <- swd99 %>% group_by(reporter, flow) %>%

# aggregate import export by country before calculating the conversion factor
# To see if the cloud looks different
swd99agg <- swd99 %>% group_by(reporter, flow, year) %>%
    summarise(weight = sum(weight),
              quantity = sum(quantity)) %>%
    mutate(conversion = weight / quantity)
ggplot(swd99agg, aes(x = conversion, y = weight)) +
  geom_point()
```

```{r conversionquantity}
p <- ggplot(filter(swd99, conversion!=Inf), 
       aes(x = conversion, y = quantity)) +
    geom_point(alpha = 1/10) +
    coord_cartesian(xlim =c(0,2000)) +
    ggtitle("Variation of the conversion factor along the quantity")
p + scale_y_log10() + ylab("Quantity (log scale)")
# p + facet_wrap(~logweight, scales="free_y")

# There seems to be flows which have the exact same quantity
# No matter the conversion factor
# This happens only for small quantities
# This is normal, because quantities are rounded at the m3
p <- ggplot(filter(swd99, conversion!=Inf & quantity<10), 
       aes(x = conversion, y = quantity)) +
    geom_point(alpha = 1/10) +
    coord_cartesian(xlim =c(0,2000)) +
    ggtitle("Variation of the conversion factor for quantities < 10m3")
p + scale_y_log10() + ylab("Quantity (log scale)")
```

```{r conversiontradevalue}
p <- ggplot(filter(swd99,conversion!=Inf), 
       aes(x = conversion, y = tradevalue)) +
    geom_point(alpha = 1/10) +
    coord_cartesian(xlim =c(0,2000)) +
    ggtitle("Variation of the conversion factor along the trade value")
p + scale_y_log10() + ylab("Tradevalue (log scale)")
```

### Conversion factors by region
```{r conversionbyregion}
swd99region<- swd99 %>% group_by(region, flow, year) %>%
    summarise(medianconv = round(median(conversion, na.rm=TRUE)),
              averageconv = round(mean(conversion, na.rm=TRUE)),
              min = min(conversion, na.rm=TRUE),
              max = max(conversion, na.rm=TRUE))
swd99region %>% select(region, year, medianconv) %>%
    dcast(region+flow~year, value.var="medianconv") %>% 
    kable(caption="Conversion factor (kg/m3) number of kg in one m3 of wood")
swd99region %>% filter(year==max(year)) %>% kable


ggplot(data=swd99region,aes(x=year, y=medianconv, color=region)) +
    geom_point() + facet_wrap(~region) + ggtitle("Conversion factor kg/m3")
```

### Conversion factors by subregion
```{r conversionbysubregion}
swd99subregion<- swd99 %>% group_by(subregion, year) %>%
    summarise(medianconv = round(median(conversion, na.rm=TRUE)))
swd99subregion %>% dcast(subregion~year) %>% 
    kable(caption="Conversion factor in kg/m3")
p <- ggplot(data=swd99subregion, aes(x=year, y=medianconv, color=subregion)) +
    geom_point() + facet_wrap(~subregion) + ggtitle("Conversion factor kg/m3")
#p
# Same plot, rescale to drop bizarre point for middle africa
p + ylim(c(0,1300))
```

