---
title: '` as.character(productcodeinreport)` discrepancies in ` reporterinreport`'
author: "Paul Rougieux"
date: "03/12/2014"
output:
  pdf_document:
    fig_height: 7
    fig_width: 10
    toc: yes
    toc_depth: 3
---

```{r packages, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
# I tried placing the yaml block after the chunks but for some reason it doesn't always generate the pdf and generates a html file instead.
# pdf is generated when the yaml block is at the top, 

library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 


```{r development, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
######### #
# ! This chunk is used for development only, remove it or change eval=FALSE.
# ! When this file is used as a template, place two  `r` in the title.
######### #
# The data frame tfdata will be passed by the report generating function to the template
# Product and reporter information can be extracted back from tfdata 
# in the template this would have to be the country appearing most in the data frame
productcodeinreport <- c(440799)
reporterinreport <- "Indonesia" 
# Test countries: Cameroon, Ghana, Indonesia
tfdata <- readdbtbl("raw_flow_yearly") %>%
    filter(productcode == productcodeinreport & 
               (reporter == reporterinreport | 
                    partner == reporterinreport)) %>%
    collect() 
```



```{r prepareproductandreporter}
numberoflargecountries <- 27
# Extract unit and check that function arguments reporterinreport and productcodeinreport are consistent with the data frame
metadata <- extractmetadata(tfdata)
stopifnot(metadata$productcode == productcodeinreport)
stopifnot(metadata$reporter == reporterinreport)
if(metadata$unit == "Volume in cubic meters") metadata$unit <- "cubic meters"

# Calculate discrepancies
tfdata  <- tfdata %>% 
    addpartnerflow %>% 
    calculatediscrepancies 

# Select the last year available for which at least half of the countries have
# quantity data 
lastyearreporter <- tfdata %>% 
    filter(reporter == reporterinreport) %>% 
    mutate(naquantity = is.na(quantity)) %>%
    group_by(year) %>%
    summarise(naratio = sum(naquantity)/n()) %>%
    filter(naratio<0.5 & year == max(year))
lastyearreporter <- lastyearreporter$year

yearinreport <- lastyearreporter

# Display number of years available only if other countries reported 
# data for more years
if (lastyearreporter != max(tfdata$year)){
    messageyears = paste("Data for other countries is available until",
                         max(tfdata$year),".")
} else {
    messageyears = ""
}
```

`r reporterinreport` reported importing from
`r length(unique(tfdata$partner[tfdata$reporter==reporterinreport & tfdata$flow=="Import"]))` countries and exporting to
`r length(unique(tfdata$partner[tfdata$reporter==reporterinreport & tfdata$flow=="Export"]))` countries 
between `r min(tfdata$year)` and `r lastyearreporter`.
`r messageyears`
 Trade value is reported in US dollars and quantity in `r metadata$unit`.
 
```{r numberofreportedflows}
tfsum <- tfdata %>% 
    mutate(isreporter = ifelse(reporter == reporterinreport,
                               "",
                               "partners")) 
# Switch import and export for the partners
tfsum[tfsum$isreporter=="partners",] <- tfsum[tfsum$isreporter=="partners",] %>%
    mutate(flow = gsub("Import","aaaaaaa",flow),
           flow = gsub("Export","Import",flow),
           flow = gsub("aaaaaaa","Export",flow))

tfsum %>%    
    group_by(year,flow, isreporter) %>%
    summarise(nrow = n(),
              value = sum(tradevalue, na.rm=TRUE),
              quantity = sum(quantity, na.rm=TRUE)) %>%
    melt(id = c("year","flow","isreporter")) %>%
    dcast(year + flow ~ isreporter + variable) %>%
    kable
```


\newpage

##  Largest trade quantities in `r yearinreport` 
 `r as.character(numberoflargecountries)` largest import or export flows reported by `r reporterinreport`. 
 Trade value is reported in US dollars and quantity in `r metadata$unit`.

```{r fig.height=14}
# Select the largest import or export flows reported by the reporter
largestreporter <- tfdata %>% 
    filterworldeu28 %>%
    filter(year == yearinreport & reporter == reporterinreport) %>% 
    arrange(desc(quantity)) %>%
    head(numberoflargecountries)

# Display latest trade flows as a table
largestreporter %>% 
    select(year, reporter, partner, flow, flag, tradevalue, quantity,
           dicrepancy_quantity = discrq) %>%
    kable

# Display these trade flows as a plot
largestreporter <- tfdata %>% 
    filterworldeu28 %>% 
    # Keep the same combination of flow / country
    filter(paste0(flow,partner) %in% 
               paste0(largestreporter$flow,
                      largestreporter$partner)) %>%
    # reshape table for use in the plot
    select(year, partner, flow, quantityreporter, quantitypartner) %>%
    arrange(-quantityreporter) %>%
    rowwise() %>%
    mutate(qmin = min(quantityreporter, quantitypartner),
           qmax = max(quantityreporter, quantitypartner)) %>%
#     rename(discrepancy = discrq) %>%
    melt(id = c("year","partner","flow", "qmin","qmax"))

ggplot(largestreporter, aes(x = year, y = value, fill = variable)) +
    geom_bar(position="dodge", stat="identity") +
    scale_fill_manual(values = c("chocolate3","black"))+
    geom_errorbar(aes(ymin = qmin, ymax = qmax, 
                      width=0.2, colour="discrepancy")) +
    scale_colour_manual(name="error bars",values="red") +
    ylab(metadata$unit) +
    facet_wrap(flow~partner, scales = "free_y", ncol = 3) +
    theme_bw() + theme(legend.position= "bottom",
                       legend.box="horizontal") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



\newpage


## Relative discrepancy
Relative discrepancy is equal to $$(quantitypartner - quantity)/ (quantity + quantitypartner)$$

```{r eval=TRUE, warning=FALSE}
## Trade flows as a point cloud
tfdata <- tfdata %>% filterworldeu28 %>% addregion 
ggplot(tfdata, aes(x = quantity, y = reldiscrq, color = regionpartner)) +
    geom_point() + facet_wrap(~flow, scales="free_x") 


ggplot(tfdata, aes(x = tradevalue, y = reldiscrv, color = regionpartner)) +
    geom_point() + facet_wrap(~flow, scales = "free_x") 

```


