```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 

```{r development, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
# Remove this chunk (or eval=FALSE) when this file is used as a template
# The data frame tfdata will be passed by the report generating function to the template
# Product and reporter information can be extracted back from tfdata 
# in the template this would have to be the country appearing most in the data frame
productcode_ <- c(440799)
reporter_ <- "Cameroon" 
tfdata <- readdbtbl("raw_flow_yearly") %>%
    filter(productcode == productcode_ & (reporter == reporter_ | partner == reporter_)) %>%
    collect
```


```{r prepareproductandreporter}
# Only the tfdata variable is passed by the function that creates the report
# Extract product code and reporter from the tfdata.
productcodeinreport <- unique(tfdata$productcode)
reporterinreport <- tfdata %>% group_by(reporter) %>% 
    summarise(nrow = n()) %>% filter(nrow == max(nrow))
reporterinreport <- reporterinreport$reporter
# Calculate discrepancies
tfdata  <- tfdata %>% addpartnerflow %>% calculatediscrepancies 
```

---
title: "440799 discrepancies in `r reporterinreport`"
author: "Paul Rougieux"
date: "03/12/2014"
output:
  pdf_document:
    toc: yes
---

## 10 largest trade flows 
### Reported by `r reporterinreport` 2 years before the last
10 largest import or export flows in  `r unique(tfdata$unit)` where `r reporterinreport` is a reporter.
```{r}
# Select the largest import or export flows reported by the reporter
largestreporter <- tfdata %>% 
    filter(year == (max(as.integer(year)) -2) & 
               reporter == reporterinreport) %>% 
    arrange(desc(quantity)) %>%
    head(10)

# Display these trade flows as a table
largestreporter %>% 
    select(year, reporter, partner, flow, flag, tradevalue, quantity) %>%
    kable
# Remove world
largestreporter <- largestreporter %>% filter(partnercode != 0)
# reshape table for use in the plot
largestreporter <- largestreporter %>%
        select(partner, weight, discrv) %>%
        arrange(-weight) %>%
        rename(discrepancy = discrv) %>%
        head(10)
```


### Reported by its trade partners 2 years before the last
10 largest import or export flows in  `r unique(tfdata$unit)` where `r reporterinreport` is a partner.
```{r}
# Select the largest import or export flows reported by the reporter
largestpartner <- tfdata %>% 
    filter(year == (max(as.integer(year)) -2) & 
               partner == reporterinreport) %>% 
    arrange(desc(quantity)) %>%
    head(10)

# Display these trade flows as a table
largestpartner %>% 
    select(year, reporter, partner, flow, flag, tradevalue, quantity) %>%
    kable
# reshape table for use in the plot
```



### Discrepancies in the raw data
```{r eval=FALSE}
# reshape the table for the plot
 dtf2 %>% 
        filter(flow == flowselected & year == max(year) & productcode==p) %>%
        select(partner, weight, discrv) %>%
        arrange(-weight) %>%
        rename(discrepancy = discrv) %>%
        head(10)

ggplot(data=largest) +
    aes(x = partner, fill = quantity) +
    geom_bar(position="dodge", stat="identity") +
    scale_fill_manual(values = c("chocolate3","black"))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle(paste(flowselected, p))
```


### Discrepancies in the refined data



