---
title: "Refining trade flows data"
author: "Paul Rougieux"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Refining trade flows data from a file or from a database}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
output:
  knitr:::html_vignette:
    toc: yes
---
```{r buildvignette, eval=FALSE, echo=FALSE}
# To build this vignette run 
devtools::build_vignettes()
# Then rebuild the package CTRL+SHIFT+B so that the vignette is 
# integrated to the package documentation
```

Function documentation can be accessed at the R 
command line by entering a question mark "?"
followed by a function name. 
For example enter `?clean` at the command prompt to see
the documentation for that function.


## Refining trade flows data from a file
```{r eval=FALSE}

```



## Refining trade flows data from a database
There are functions to clean (estimate missing quantities) trade flow data and write this cleaned data into the database.

The methodology report explains the data 
refining process in details.
The source code of the `clean()` fonction gives details on the implementation of the data refining workflow.
```{r eval=FALSE}
?clean # View help for this function
sawnwood99_raw <- readdbproduct(440799, "raw_flow_yearly")
sawnwood99_validated <- clean(sawnwood99_raw)

# Those 2 functions together in one go
cleandbproduct(440799, "raw_flow_yearly", "validated_flow_yearly")
```


## Notes on partner flow
The parter flow in the database is the one before the modification.
