---
title: "Installation and configuration"
author: "Paul Rougieux"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Installation and configuration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
output:
  knitr:::html_vignette:
    toc: yes
---
```{r eval=FALSE, echo=FALSE}
# To build this vignette run 
devtools::build_vignettes()
# Then rebuild the package CTRL+SHIFT+B so that the vignette is 
# integrated to the package documentation
```
 
 
## Installation
This package can be installed with the devtools package.
```{r eval=FALSE}
library(devtools)
install_bitbucket("paul4forest/tradeflows")
```

### Devtools package installation
If the devtools package is not installed on your system.
```{r eval=FALSE}
install.packages("devtools")
```
You may need to upgrade the libcurl library on your system's command line
```
sudo apt-get install libcurl4-gnutls-dev 
```


## Configuration
Database configuration file and column names are located under the local package installation directory. 
This directory is shown by the R command:
```{r eval=FALSE}
system.file("config", package="tradeflows")
```
Database connection parameters can be entered
in the databaseconfig.R file.

Load the package and use the database configuration
with the commands:
```{r eval=FALSE}
library(tradeflows)
setdatabaseconfig(reload=TRUE)
```
The database configuration file can be edited and reloaded.
`setdatabaseconfig(reload=TRUE)` displays the location of the database configuration file.


The configuration table columnnames located in config/column_names.csv
contains columns specifying which efi 
columns names are used in the trade flows database
raw_flow and validated_flow tables.


### Digging into the code
Start by looking at files in the R/ directory

* database.R is doing the database interaction
* clean.R is cleaning the datai


The function cleandb() will feed data into the database table(s) validated_flow
updates will be done on a product basis, at the 6 digit level. The cleaning script will:

1. Delete all flows for a product
  (between all reporter and partner countries in all years),
2. Enter All validated flows for that product.

The clean() function can also run from a bash shell
(your system's command line)
```
Rscript -e 'library(tradeflows)' -e 'cleandbproduct(440799, tableread =  "raw_flow_yearly", tablewrite = "validated_flow_yearly")
```
