---
title: "Loading and refining trade flows data from the Comtrade API"
author: "Paul Rougieux"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Loading and refining trade flows}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
output:
  knitr:::html_vignette:
    toc: yes
---
```{r buildvignette, eval=FALSE, echo=FALSE}
# To build this vignette run 
devtools::build_vignettes()
# Then rebuild the package CTRL+SHIFT+B so that the vignette is 
# integrated to the package documentation
```

Function documentation can be accessed at the R 
command line by entering a question mark "?"
followed by a function name. 
For example enter `?loadcomtradebycode` at the command prompt to see
the documentation for that function.


## Loading Comtrade data in 2 stages
### Stage 1 load from one reporting country 
Comtrade API documentation is available at http://comtrade.un.org/data/doc/api/

The function `loadcomtradebycode()` loads a JSON file 
from the comtrade API in a temporary directory and returns a dataframe.
It loads trade flows between one reporting country (column `reporter`) and all its partner countries (column `partner`).
```{r loadcomtradebycode, echo=TRUE, eval=TRUE}
library(tradeflows)
library(knitr) # Package to display nice tables

# Load "other sawnwood" flows reported by France (country code 251) 
# for the last 5 years available in comtrade "recent"
other_sawnwood_fr <- loadcomtradebycode(440799, 251, "recent")
# Rename columns so that they are more readable
other_sawnwood_fr <- renamecolumns(other_sawnwood_fr)
other_sawnwood_fr %>% 
    # Select only some of the column so that the table fits on the page
    select(year, reporter, partner, flow, productcode,
           unit, quantity, weight, tradevalue, flag) %>%
    head %>%
    # Show a nice table
    kable    
```

### Stage 2 load from all reporting countries
The comtrade database API has limitation of 5 countries per call.
There are approximately 250 countries in the database.
The function `loadcomtradewithpause` loads all world trade flows
for one product. It sends several function calls to the API, in batches of 3 countries, then combines the downloaded data in one data frame.

The Comtrade API has a limitation of 100 calls per hour.
Given a list of products, loadcomtradewithpause() 
will wait one hour before downloading the next product.

```{r loadcomtradewithpause1, eval=FALSE}
loadcomtradewithpause(440799, year = "recent", path = "data-raw/")
```

RData is stored in the  `path` directory given above.
A csv file can also be saved in order to visualise raw data in a spreadsheet. 
```{r loadcomtradewithpausecsv, eval=FALSE}
loadcomtradewithpause(440799, year = "recent", path = "data-raw/",
                      writecsv = TRUE)
```

The instruction below loads all sawnwood products. It will last 14 hours because it's waiting one hour between each product. 
```{r loadcomtradewithpausemany, eval=FALSE}
# Find sawnwood product codes from the itto classification
sawnwoodcodes <- filter(classificationitto, 
                        nomenclature == "HS12",
                        product == "SAWNWOOD")$productcodecomtrade
loadcomtradewithpause(sawnwoodcodes, year = "recent", 
                      path = "data-raw/")
```

In case you are launching this lengthy instruction 
on a remote linux server,
it might be usefull to launch it from within a 
[GNU-Screen](http://www.gnu.org/software/screen/manual/screen.html) 
session. To be able to disconnect a reconnect to it at a later stage.


## Using a MySQL database
Optionally, data can be read from a MySQL database. 
The function `readdbproduct` loads raw data from a 
MySQL database table:
```{r readdbproduct, eval=FALSE}
sawnwood99 <- readdbproduct(440799, "raw_flow_yearly")
```

The function  readdbtbl returns a `dplyr` `tbl` objects
which can be used with dplyr verbs to perform data manipulations.
```{r readdbtbl, eval=FALSE}
# Five largest log trade flows in value in 2012
# Find logs product codes from the itto classification
roundwoodcodes <- filter(classificationitto, 
                        nomenclature == "HS12",
                        product == "LOGS")$productcodecomtrade

readdbtbl("raw_flow_yearly") %>%
    filter(productcode %in% roundwoodcodes &
               year == 2012 &
               partnercode != 0) %>% # Remove world aggregates
    arrange(desc(tradevalue)) %>% # Sort by trade value
    select(year, reporter, partner, flow, productcode, tradevalue) %>%
    head() %>%
    kable
```

```{r displaycacheversionoftable, echo=FALSE}
# Display this table also on systems where the vignette cannot be 
# created
logs2012 <- structure(list(year = c("2012", "2012", "2012", "2012", "2012", "2012"), reporter = c("China", "China", "Russian Federation", 
"New Zealand", "China", "USA"), partner = c("Russian Federation", 
"New Zealand", "China", "China", "USA", "China"), flow = c("Import", 
"Import", "Export", "Export", "Import", "Export"), productcode = c(440320, 440320, 440320, 440320, 440320, 440320), tradevalue = c(1394408301, 1103088193, 1000281108, 847521843, 601848451, 569742376)), .Names = c("year", "reporter", "partner", "flow", "productcode", "tradevalue"), row.names = c(NA, -6L), class = "data.frame")
logs2012 %>% kable
```


The project in which this package was created 
used external scripts - in the PHP programming language -
to load data from comtrade and write it into the database. 
Although this is feasible, the package currently
doesn't implement writing raw trade flows data 
into the database.

However, there are functions to clean (estimate missing quantities) trade flow data and write this cleaned data into the database.
 

## Refining trade flows data
The methodology report explains the data 
refining process in details.
The source code of the `clean()` fonction gives details on the implementation of the data refining workflow.
```{r eval=FALSE}
?clean # View help for this function
refined_data <- clean(raw_data)
```
