---
title: "Network visualisation of forest products trade flows"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---



```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=TRUE)
library(tradeflows)
library(dplyr)
library(igraph)
``` 

```{r load}
# This chunk can be deleted once this file is converted to a template
load("440799.RData")
swd99 <- renamecolumns(dtf, "comtrade", "efi")
```
## Load data
This report is generated automatically based on data from the 
[UN Comtrade API](http://comtrade.un.org/data/Doc/api/ex).
Tables and graphs are in development.
They are intended for internal use within the TradeFLowsDB project.

### Warning for the igraph package
* Column name "weight" in the context of the igraph package
* It has to have a name column
* Coordinates has to be with x and y
* Is has to have a "name" column which are these ISO3 codes in the attributes object.

### Warning for unavailable iso codes
Areas, nes, Free Zones | Other Asia, nes | North America and Central America, nes,
Other Europe, nes |Other Africa, nes     

See table below. list of partner for which  unavailable iso codes
```{r}
swd99 %>% filter(is.na(partneriso)) %>%
    select(partner, partneriso) %>%
    unique
```

## Prepare data
```{r prepare}
# reporter2exclude <- reportercomtrade %>% filter(is.na(region))
#     filter(reportercode %in% reporter2exclude$reportercode |
#                partnercode %in% reporter2exclude$reportercode )

swd99 <- renamecolumns(dtf, "comtrade", "efi") %>%
    filter(flow == "Import" & year ==2011) %>%
    # Remove World and EU28 
    filter(!reportercode %in% c(0,97) &
               !partnercode %in% c(0,97)) %>%
    # Remove those trade flows from a country with itself
    filter(reportercode != partnercode) %>%
    # Remove reporteriso and partneriso which are not available
    filter(!is.na(reporteriso) & !is.na(partneriso))

data_frame <- swd99 %>% 
  select(EXPORTER = partneriso,
         IMPORTER = reporteriso,
         weight = tradevalue) %>%
  mutate(weight = round(weight/1000))
# show begining of table
kable(head(data_frame))
```

### Prepare attributes tables and Generate network visualisation
```{r}
# Sort all countries by alphabetical order
attributes_COO <- data.frame(name = unique(c(data_frame$EXPORTER, data_frame$IMPORTER))) %>%
    arrange(name) %>%
    mutate(name = as.character(name))

############################################### #
# Experimental, comment this out if not needed  #
############################################### #
# Add all combination of EXPORTER and IMPORTER codes 
# to the data frame to try to avoid the error message "vertex names in edge list..."
# When running the graph.data.frame function
# > data_graph <- graph.data.frame(d = data_frame,   vertices = attributes_COO)
# Error in graph.data.frame(d = data_frame, vertices = attributes_COO) : 
#   Some vertex names in edge list are not listed in vertex data frame
dtf <- data.frame()
for (v in attributes_COO$name){
    dtf <- rbind(dtf,
                 data.frame(EXPORTER=v, IMPORTER=attributes_COO$name))
}
data_frame <- merge(data_frame, dtf, all=TRUE) 
data_frame$weight[is.na(data_frame$weight)] <- 0
###################################

attributes_COO$id <- row.names(attributes_COO)
# Sum of export for each country
exports <- data_frame %>% group_by(EXPORTER) %>%
    summarise(x = sum(weight)) %>% rename(name = EXPORTER)
# Sum of import for each country
imports <- data_frame %>% group_by(IMPORTER) %>%
    summarise(y = sum(weight)) %>% rename(name = IMPORTER)
# Merve sum of export and sum of import into the attributes table
attributes_COO <- attributes_COO %>%
    merge(exports) %>% 
    merge(imports) %>%
    mutate(Tot.tr = x + y, 
           size = Tot.tr / max(Tot.tr)*100, # Total trade normalised from 0 to 100
           percent = Tot.tr / sum(Tot.tr)*100) 

# Rename this column percent to "%" very bad practice
names(attributes_COO)[names(attributes_COO)=="percent"] <- "%"
# show begining of table
kable(head(attributes_COO))

################################ #
# Generate network visualisation #
################################ #
options(scipen = 3)

data_graph <- graph.data.frame(data_frame) 
data_graph1 <- delete.edges(data_graph,   E(data_graph)[ weight < 479]    )

# attributes_COO <- read.csv("docs/development/networkvisualisation/Attributes COO.csv", header=TRUE)
# attributes_COO <-  attributes_COO %>% filter(!name %in% data_frame$EXPORTER)
# attributes_COO %>% filter(!name %in% data_frame$IMPORTER)

# attributes_COO = cbind(1:length(attributes_COO [,1]), attributes_COO)
data_graph <- graph.data.frame(d = data_frame, vertices = attributes_COO)

# Specify the layout
Layout_auto <- layout.auto(data_graph1, dim=2)
V(data_graph1)$size=degree(data_graph)*200000
E(data_graph1)[ weight > 3147 ]$color <- " grey "
E(data_graph1)[ weight > 552782 ]$color <- "red"

plot(data_graph1,
     layout= Layout_auto, rescale=FALSE, 
     edge.arrow.size=0.7, edge.width=E(data_graph1)$weight/350000,
      xlim=range(attributes_COO[,4]), ylim=range(attributes_COO[,5]), axes=TRUE, 
#       xlim=round(range(attributes_COO[,4])/1000), ylim=range(attributes_COO[,5]), axes=TRUE, 
     xlab="Export (1000 USD)", ylab="Import (1000 USD)") 
```


### Error
"Some vertex names in edge list are not listed in vertex data frame"
```{r errorvertexnames}
attributes_COO$name %in% data_graph1
str(data_graph1)

# Do we loose countries?
sum(!attributes_COO$name %in% data_frame$EXPORTER)
attributes_COO$name[!attributes_COO$name %in% data_frame$EXPORTER]
sum(!data_frame$EXPORTER %in% attributes_COO$name)
sum(!attributes_COO$name %in% data_frame$IMPORTER)
sum(duplicated(data_frame[c("EXPORTER","IMPORTER")]))

sum(!data_frame$EXPORTER %in% dtf$EXPORTER)
unique(data_frame$EXPORTER[!data_frame$EXPORTER %in% dtf$EXPORTER])
sum(!data_frame$IMPORTER %in% dtf$IMPORTER)

```

