---
title: "Network visualisation of forest products trade flows"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---



```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE)
library(tradeflows)   # Which package is this?
library(dplyr)
library(igraph)
library(caTools)
library(maps)  
library(mapdata) 
``` 

```{r load}
# This chunk can be deleted once this file is converted to a template
load("data-raw/comtrade/440799.RData")
swd99 <- renamecolumns(dtf, "comtrade", "efi")
```
## Load data
This report is generated automatically based on data from the 
[UN Comtrade API](http://comtrade.un.org/data/Doc/api/ex).
Tables and graphs are in development.
They are intended for internal use within the TradeFLowsDB project.

### Warning for the igraph package
* Column name "weight" in the context of the igraph package
* Coordinates has to be with x and y
* Is has to have a "name" column which are these ISO3 codes in the attributes object.

### Warning for unavailable iso codes
Areas, nes, Free Zones | Other Asia, nes | North America and Central America, nes,
Other Europe, nes |Other Africa, nes     

See table below. list of partner for which  unavailable iso codes
```{r}
swd99 %>% filter(is.na(partneriso)) %>%
    select(partner, partneriso) %>%
    unique
```

## Prepare data
```{r prepare}
# reporter2exclude <- reportercomtrade %>% filter(is.na(region))
#     filter(reportercode %in% reporter2exclude$reportercode |
#                partnercode %in% reporter2exclude$reportercode )

swd99 <- renamecolumns(dtf, "comtrade", "efi") %>%
    filter(flow == "Import" & year ==2011) %>%
    # Remove World and EU28     !!!! Year should not be 2011 fixed, but last one -2!!!!!
    filter(!reportercode %in% c(0,97) &
               !partnercode %in% c(0,97)) %>%
    # Remove those trade flows from a country with itself   !!!!! THIS DID NOT WORK !!!!!!!!! 
    filter(reportercode != partnercode) %>%
    # Remove reporteriso and partneriso which are not available
    filter(!is.na(reporteriso) & !is.na(partneriso))

data_frame <- swd99 %>% 
  select(EXPORTER = partneriso,
         IMPORTER = reporteriso,
         weight = tradevalue) %>%
  mutate(weight = round(weight/1000))
# show begining of table
kable(head(data_frame))
```

### Prepare attributes tables and Generate network visualisation
```{r}
# Sort all countries by alphabetical order
attributes_COO <- data.frame(name = unique(c(data_frame$EXPORTER, data_frame$IMPORTER))) %>%
    arrange(name) %>%
    mutate(name = as.character(name))

############################################### #
# Experimental, comment this out if not needed  #
############################################### #
# Add all combination of EXPORTER and IMPORTER codes 
# to the data frame to try to avoid the error message "vertex names in edge list..."
# When running the graph.data.frame function
# Error in graph.data.frame(d = data_frame, vertices = attributes_COO) : 
#   Some vertex names in edge list are not listed in vertex data frame
# dtf <- data.frame()
# for (v in attributes_COO$name){
#     dtf <- rbind(dtf,
#                  data.frame(EXPORTER=v, IMPORTER=attributes_COO$name))
# }
# data_frame <- merge(data_frame, dtf, all=TRUE) 
# data_frame$weight[is.na(data_frame$weight)] <- 0
# ###################################

attributes_COO$id <- row.names(attributes_COO)
# Sum of export for each country
exports <- data_frame %>% group_by(EXPORTER) %>%
    summarise(x = sum(weight)) %>% rename(name = EXPORTER)
# Sum of import for each country
imports <- data_frame %>% group_by(IMPORTER) %>%
    summarise(y = sum(weight)) %>% rename(name = IMPORTER)
# Merve sum of export and sum of import into the attributes table   !!! The command below decreases the number of rows in attrib.!
attributes_COO <- attributes_COO %>%
    merge(exports, all.x=TRUE) %>% 
    merge(imports, all.x=TRUE)
attributes_COO$x[is.na(attributes_COO$x)] <- 0
attributes_COO$y[is.na(attributes_COO$y)] <- 0
attributes_COO <- attributes_COO %>%
  mutate(Tot.tr = x + y, 
           size = Tot.tr / max(Tot.tr)*100, 
           percent = Tot.tr / sum(Tot.tr)*100) 
# Total trade normalised from 0 to 100

# This creates the data graph file with linked attributes file!
data_graph <- graph.data.frame(d = data_frame, vertices = attributes_COO)


# Rename this column percent to "%" very bad practice  # perhaps we do not need to do this
# names(attributes_COO)[names(attributes_COO)=="percent"] <- "%"
# show begining of table
kable(head(attributes_COO))

################################ #
# Generate network visualisation #
################################ #
# this is a penalty for using scientific notations
options(scipen = 3)    
# This creates an graph object 
# data_graph <- graph.data.frame(data_frame) 

# attributes_COO <- cbind(1:length(attributes_COO [,1]), attributes_COO)
# This deletes ties (trade flows) below the mean value of trade flow.  It is better to tye to do this just under plot! 
# data_graph1 <- delete.edges(data_graph,   E(data_graph)[ weight < 489]    ) 
# attributes_COO <- read.csv("docs/development/networkvisualisation/Attributes COO.csv", header=TRUE)
# attributes_COO <-  attributes_COO %>% filter(!name %in% data_frame$EXPORTER)
# attributes_COO %>% filter(!name %in% data_frame$IMPORTER)


# data_graph <- graph.data.frame(d = data_frame, vertices = attributes_COO)
# pdf("Graph_1.pdf")
# Specify the layout
Layout_auto <- layout.auto(data_graph, dim=2)
# Specify the color for trade flows above the mean 

# This automatically defines red color for top 10 trade flows, other ones are grey

E(data_graph)[ weight > sort(data_frame[,3], TRUE)[10] ]$color <- "red"
E(data_graph)[ weight < sort(data_frame[,3], TRUE)[10] ]$color <- " grey "


# This plots the graph, where x is export, y is import axis
plot(data_graph,layout= Layout_auto, vertex.size=V(data_graph)$size/3, edge.arrow.size=0.7, 
     edge.width=E(data_graph)$weight/150000, rescale=FALSE, xlim=range(attributes_COO[,3]), 
     ylim=range(attributes_COO[,4]), axes=TRUE, xlab="Export (1000 USD)", ylab="Import (1000 USD)" )   

# dev.off()    







     
```

```{r errorvertexnames0}
# CONCENTRIC CIRCLES VISUALIZATION - WE NEED THE CIRCLE CALCULATION FROM EXCEL!!!


```


```{r errorvertexnames}
# WORLD MAP VISUALIZATION
# This is done through a prepared blank world map (wold-map.gif), and a file that contains coordinates for each
# coutry (World Map COO.csv). 

# Making of new attributes_COO which do not have coordinates
attributes_COO1 <- attributes_COO
attributes_COO1$x <- NULL
attributes_COO1$y <- NULL

# Read attributes map all.csv
attributes_map_all <- read.csv("data-raw/World Map COO.csv", header=TRUE)

# Merge attributes_COO1 and attributes_map_all to a new attribute object which has all needed country # ISO 3 codes and coordinates from the map of the world 
# Coordinates of countries correspond to pixel coordinates of the world map
# Y coordinates are calculated by 2526-y. This is done within the file itself, no need to do anything # here
attributes_WM <- merge(attributes_COO1, attributes_map_all, by="name")
# keep only unique entries
attributes_WM <- attributes_WM[!duplicated(attributes_WM[,1]),]
# change column name 'name' to 'label'
colnames(attributes_WM)[1] <- "label"


# Make a graph object                        
data_graph_WM <- graph.data.frame(d = data_frame, vertices = attributes_WM)

# E(data_graph_WM)[ weight > sort(data_graph_WM[,3], TRUE)[10] ]$color <- "red"
# E(data_graph_WM)[ weight < sort(data_graph_WM[,3], TRUE)[10] ]$color <- "grey"
# W_10 <- sort(data_frame[,3], TRUE)[10] 
# data_graph_WM <- delete.edges(data_graph_WM, E(data_graph_WM) [ weight < "W_10" ])

# This leaves only top 10 trade flows
data_graph_WM <- delete.edges(data_graph_WM, E(data_graph_WM) [ weight < sort(data_frame[,3], TRUE)[10] ] )

E(data_graph_WM)[ weight > sort(data_frame[,3], TRUE)[10] ]$color <- "red"

Layout_auto_WM <- layout.auto(data_graph_WM, dim=2)


# data_graph_WM <- delete.vertices(data_graph_WM,which(degree(data_graph_WM)<1))
# This adds a column in attributes that represents degree 
attributes_WM$degree      <- degree(data_graph_WM, mode = c("all", "out", "in", "total"))
# This sets degree of all connected vertexes to 1
attributes_WM$degree[attributes_WM$degree >=1] <- 1 
# This sets vertex label to 0 for all those with degree 0
attributes_WM$label[which(attributes_WM$degree == 0)] <- 0
# Change 0 vertex names to NA 
attributes_WM[,1][which(attributes_WM[,1]<1)] <- NA









# par(pin=c(20, 10.104))


      # plot similar to E/I, but with a fixed aspect ratio, i.e. asp=y/x, and labels are drawn from attributes
plot(data_graph_WM,layout= Layout_auto_WM, vertex.size=V(data_graph_WM)$size/8, edge.arrow.size=0.7, 
     edge.width=3, asp=0.5052, vertex.label=attributes_WM[,1] )




# xlim=range(1, 5000), ylim=range(1, 2526)

# MARGINS - A vector of the form c(bottom, left, top, right)  in plot margin=c(-0.15,0,-0.15,-0.15)

##############
# Map instruction
map('worldHires', asp=0.5052)
# To be able to stack chart on map
par(new=TRUE)

# Plotting instruction
plot(data_graph_WM,layout= Layout_auto_WM, 
     vertex.size=V(data_graph_WM)$size/8, edge.arrow.size=0.7, 
     edge.width=3, vertex.label=attributes_WM[,1], 
      asp=0.4)
#       asp=0.5052) 


# Without aspect ratio
# Map instruction
map('worldHires')
# To be able to stack chart on map
par(new=TRUE)

# Plotting instruction
plot(data_graph_WM, layout= Layout_auto_WM, 
     vertex.size=V(data_graph_WM)$size/8, edge.arrow.size=0.7, 
     edge.width=3, vertex.label=attributes_WM[,1])

# Try to plot with ggplot, 
# here are the name of countries using your coordinate system
# Do not use this example
ggplot(attributes_map_all, aes(x,y, label=name)) + 
    geom_point(size=20,colour="lightgrey") + geom_text() +
    theme_bw()

# create a layer of borders
mapWorld <- borders("world", colour="gray50", fill="gray50") 
mp <- ggplot() + mapWorld
# Use the list of centre point code in longitude latitude
# To add centre of countries with name


###########UNTIL HERE###################
```



```{r tryingggplot}
library(ggmap)
library(maptools)
library(maps)
 
 
visited <- c("SFO", "Chennai", "London", "Melbourne", "Johannesbury, SA")
ll.visited <- geocode(visited)
visit.x <- ll.visited$lon
visit.y <- ll.visited$lat 
#Using GGPLOT, plot the Base World Map
mp <- NULL
mapWorld <- borders("world", colour="gray50", fill="gray50") # create a layer of borders
mp <- ggplot() + mapWorld
 
#Now Layer the cities on top
mp <- mp + geom_point(aes(x=visit.x, y=visit.y) ,color="blue", size=3)
mp 


``` 


```{r eval=FALSE}
par(fig=c(0.1, 1, 0, 0.9)
opar <- par(fin=c(4,2), new=TRUE )
par(mar = rep(0, 4))
map('worldHires', asp=0.5052, mar=c(2,2,2,2) )
plot(data_graph_WM,layout= Layout_auto_WM, vertex.size=V(data_graph_WM)$size/8, edge.arrow.size=0.7, 
     edge.width=3, vertex.label=attributes_WM[,1], asp=0.5052, margin=c(-0.2,-0.2,-0.2,-0.2))

par(opar)
# The scale of the trade plot and the map is not adequate, the scale of x,y in plot of trade is too small
# we need to increase 

par(plt=c(0.75, 0.95, 0.75, 0.95), new=TRUE)

par(new=TRUE)

# This links the map and the trade graph into a same plot. Fig also sets the 'area' where to plot the trade network. 
# Four elements within Fig are range of x and y plot area. 
 par(fig=c(0.05, 1, 0, 0.9), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)

 par(fig=c(0.05, 1, 0, 0.9), oma=c(0, 0, 0, 0),  new=TRUE)

map("worldHires", border=0, ylim=c(-80, 80), fill=TRUE, 
      col="white",mar=rep(0,4))




#######################


WM <-  read.gif("World_map.gif", flip=FALSE)
n <- dim(WM$image)
image(t(WM$image)[n[2]:1,n[1]:1],col=WM$col,axes=F)
WM$col[WM$col=="#000000FF"] <- "#FFFFFFFF"
op <- par(new=T)
plot(data_graph_WM,layout= Layout_auto_WM, vertex.size=V(data_graph_WM)$size/6, edge.arrow.size=0.7, 
     edge.width=3, asp=0.5052, vertex.label=attributes_WM[,1], new=T )
par(op)








###  This is by usage of caTools ### Does not work.... black is white and vice versa!!
WM <-  read.gif("World_map.gif")
n <- dim(WM$image)
image(t(WM$image)[n[2]:1,n[1]:1],axes=F)
### Changes black to white.... but still lines are double, this is a dead end
WM$col[WM$col=="#000000FF"] <- "#FFFFFFFF"
n <- dim(WM$image)
op <- par(new=T)
plot(data_graph_WM,layout= Layout_auto_WM, vertex.size=V(data_graph_WM)$size/6, edge.arrow.size=0.7, 
     edge.width=3, asp=0.5052, vertex.label=attributes_WM[,1] )
par(op)

image(t(Gif$image)[n[2]:1,n[1]:1],col=Gif$col,axes=F)
op <- par(new=T)
plot(1:100,new=T)
par(op)

###############################


Gif <- read.gif("World_map.gif")

n <- dim(Gif$image)
image(t(Gif$image)[n[2]:1,n[1]:1],col=Gif$col,axes=F)
op <- par(new=T)
plot(1:100,new=T)
par(op)








df$Items[which(df$Store.Type == "A" | df$Store.Type == "C" )] <- 0

attributes_WM$name <- with(attributes_WM, if(degree == 0)name <-NA)



which(V(data_graph_WM)$name [degree>0] )
V(data_graph_WM)$name = degree(data_graph_WM)
  
  degree < W_10
V(net)$name   vertex.label=V(net)$name 
in plot               vertex.label=V(data_graph_WM)$label=="USA"
V(data_graph_WM)[degree>0]

V(data_graph_WM)$label.cex <- attributes_WM$degree-5

V(data_graph_WM)$label.cex = attributes_WM$degree


V(data_graph_WM)$label.cex = 0

V(g1)$label.cex[1:20 %% 2 == 0] = 0.

V(net)$name 

V(data_graph_WM)$name=="USA"


plot(data_graph_WM,layout= Layout_auto_WM, vertex.size=V(data_graph_WM)$size/8, edge.arrow.size=0.7, 
     edge.width=3, asp=0.5052,  label.cex=attributes_WM$degree )
     
     
     vertex.label=V(data_graph_WM)$name=="USA")


ifelse(attributes_WM$degree = 1, vertex.label=V(data_graph_WM)$name, V(data_graph_WM)$name=NA )

data_graph_WM=delete.vertices(data_graph_WM,which(degree(data_graph_WM)<1))


TRY WITH MERGE!!!!!


```







### Error
"Some vertex names in edge list are not listed in vertex data frame"
```{r errorvertexnames}
attributes_COO$name %in% data_graph1
str(data_graph1)

# Do we loose countries?
sum(!attributes_COO$name %in% data_frame$EXPORTER)
attributes_COO$name[!attributes_COO$name %in% data_frame$EXPORTER]
sum(!data_frame$EXPORTER %in% attributes_COO$name)
sum(!attributes_COO$name %in% data_frame$IMPORTER)
sum(duplicated(data_frame[c("EXPORTER","IMPORTER")]))

sum(!data_frame$EXPORTER %in% dtf$EXPORTER)
unique(data_frame$EXPORTER[!data_frame$EXPORTER %in% dtf$EXPORTER])
sum(!data_frame$IMPORTER %in% dtf$IMPORTER)

```

