---
title: "Flags in the validated dataset"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tradeflows)
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
opts_knit$set(fig.width=12)
library(dplyr)
library(ggplot2)
library(tidyr)

``` 


# Warning!!!  number of flags are biaised if you don't remove products at the 4 digit 
level from the database. Start with the last chunk and redo this flag extraction once and 
for all. 

# Flag use in number of flows

```{r}
flagn <- readdbtbl("validated_flow_yearly") %>%
    group_by(flag) %>%
    summarise(number_of_flows = n()) %>%
    collect() %>%
    mutate(flag = as.factor(flag))
ggplot(flagn, aes(x = flag, y = number_of_flows)) + geom_bar(stat  = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


# Flag use in volume
Some flows are expressed in cubic meters, others in kg and others in number of units.
The graph below illustrates the distribution of flags for the various units.
```{r}
flagq <- readdbtbl("validated_flow_yearly") %>%
    group_by(flag, unit) %>%
    summarise(quantity = sum(quantity)) %>%
    collect() %>% ungroup() %>%
    mutate(flag = as.factor(flag))
ggplot(flagq, aes(x = flag, y = quantity)) + geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_grid(unit~., scales="free_y")
```


# Number of flags by product 
Simplify to the 2 digit level
```{r}
flagpn <- readdbtbl("validated_flow_yearly") %>%
    group_by(flag, productcode) %>%
    summarise(number_of_flows = n()) %>%
    collect() %>% ungroup() %>%
    mutate(flag = as.factor(flag),
           digits = round(productcode/1e4)) %>%
    filter(digits > 0)

ggplot(flagpn, aes(x = flag, y = number_of_flows)) + geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_grid(digits~., scales="free_y")
```
