---
title: "Import Export mismatch"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tradeflows)
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(dplyr)
library(ggplot2)
library(tidyr)
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
``` 


# Introduction / Description of the issue

Since the update in comext data structure, unitcode are now part of the main table
"raw_comext_monthly". It is thus not necessary anymore to add unit codes to the main table.
The `cleancomext()` function needs to be updated to reflect that.
 

In 2018 there was an update to the Comext data structure.
The update requires a modification of the database structure and of
the way units are dealt with in the cleaning functions. 
The issue lies in the `loadcomext1product()` function, 
itself calling the `eutradeflows::addunit2tbl()` function.
The table already has a unit code, we only need to add the unit description. 

```{r}
# Output of an erroneous run before the change
cleancomext("tradeflows")
#
# Starting to clean on 2018.06.29 at 09:52
# 
# Loading table definitions from:
# /home/debian/R/x86_64-pc-linux-gnu-library/3.4/eutradeflows/config/vld_comext.sql
# into the `tradeflows` database.
# ERROR 1060 (42S21) at line 104: Duplicate column name 'unitcode'
```

# Reproduce the bug locally.

## Missing data in the raw codes table
The lack of data inthe raw codes tables has to be addressed in the harvester package. 
See ~/R/tradeharvester/docs/harvest.Rmd for details.

## Get a backup of the raw database table
See also backup of the validated data below
```{r}
# Dump codes tables
eutradeflows::dumptable("tradeflows", "raw_comext_partner", "/mnt/sdb/public/sqldump/")
eutradeflows::dumptable("tradeflows", "raw_comext_product", "/mnt/sdb/public/sqldump/")
eutradeflows::dumptable("tradeflows", "raw_comext_reporter", "/mnt/sdb/public/sqldump/")
eutradeflows::dumptable("tradeflows", "raw_comext_unit", "/mnt/sdb/public/sqldump/")
eutradeflows::dumptable("tradeflows", "raw_comext_unit_description", "/mnt/sdb/public/sqldump/")
# Dump main table
eutradeflows::dumptable("tradeflows", "raw_comext_monthly", "/mnt/sdb/public/sqldump/")
# The main table takes a long time try to load data for one product instead

# load data for one product only
# Dump by queru SO question
# export from the Mysql CLI
# $ mysql -e "select * from raw_comext_monthly where productcode = 44071091" tradeflows > swd44071091.sql
# The result is not an SQL file
# Export using the mysqldump tool
# mysqldump --databases tradeflows --tables raw_comext_monthly --where="productcode = 44071091" > swd44071091.sql
```


## Load the dumps into the local database on my laptop

```{r}
library(eutradeflows)
# Codes tables
loadtabledump("tradeflows", "~/downloads/raw_comext_partner.sql.7z")
loadtabledump("tradeflows", "~/downloads/raw_comext_product.sql.7z")
loadtabledump("tradeflows", "~/downloads/raw_comext_reporter.sql.7z")
loadtabledump("tradeflows", "~/downloads/raw_comext_unit.sql.7z")
loadtabledump("tradeflows", "~/downloads/raw_comext_unit_description.sql.7z")
# Main table
loadtabledump("tradeflows", "~/downloads/raw_comext_monthly.sql.7z")
```

Now the raw data is in my local database and I can reproduct the error.
```{r}
cleancomext("tradeflows")
# Recreate the error with a function further down the nest 
dbname = "tradeflows"
rawtabletemplate = "raw_comext_monthly_template"
vldtabletemplate = "vld_comext_monthly_template"
tablewrite = "vld_comext_monthly"
tablepriceconversion = "vld_comext_priceconversion"
templatecharacters = "template"
logfile = paste0('~/public_html/log/validate',format(Sys.Date(), '%Y'),'.txt')

# Now replace all RMySQL connection objects with  RMariaDB
cleancomextmonthly1product(RMariaDBcon = RMariaDBcon,
                           productanalysed = productcode,
                           tablearchive = tablearchive,
                           tablerecent = tablerecent,
                           tablewrite = tablewrite,
                           tablepriceconversion = tablepriceconversion)
```


## Backup of the validated data
```{r}
sqldumpfolder <- "/mnt/sdb/public/sqldump/"
eutradeflows::dumptable("tradeflows", "vld_comext_partner", sqldumpfolder)
eutradeflows::dumptable("tradeflows", "vld_comext_product", sqldumpfolder)
eutradeflows::dumptable("tradeflows", "vld_comext_reporter", sqldumpfolder)
eutradeflows::dumptable("tradeflows", "vld_comext_unit", sqldumpfolder)
eutradeflows::dumptable("tradeflows", "vld_comext_unit_description", sqldumpfolder)
# Dump main table
eutradeflows::dumptable("tradeflows", "vld_comext_monthly", sqldumpfolder)
```




## Sample code
```{r}
# Connect to the database
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
# Load data for product 44079910
# The issue is within that code:
dtf <- loadcomext1product(con,
                          productanalysed = "44071091",
                          table = "raw_comext_monthly")
# which itself calls `eutradeflows::addunit2tbl`
# The problem is that the unit code is already there. 

head(dtf)
unique(dtf$year)
```

## Investigate

```{r}
productanalysed <- "44071091"
dtfr <- tbl(con, "raw_comext_monthly") %>%
    filter(productcode == productanalysed) %>%
    collect()
```


### Load missing metadata
Units seem to be missing from the database. 
Trying to reload them directly on the server with the R harvest function below
```
# I copied the following R commants to the server directly 
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
tradeharvester::harvestcomextmetadata(RMariaDBcon=con, rawdatafolder='/mnt/sdb/public', pause=0)
```

It worked, then dump the corresponding metadata tables with:
```{r}
dumptable("tradeflows", "raw_comext_partner", "/mnt/sdb/public/sqldump/")
dumptable("tradeflows", "raw_comext_product", "/mnt/sdb/public/sqldump/")
dumptable("tradeflows", "raw_comext_reporter", "/mnt/sdb/public/sqldump/")
dumptable("tradeflows", "raw_comext_unit", "/mnt/sdb/public/sqldump/")
dumptable("tradeflows", "raw_comext_unit_description", "/mnt/sdb/public/sqldump/")
```



```{r dbDisconnect}
RMariaDB::dbDisconnect(con)
```


