---
title: "List of reports for the Virtual Private Server"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tradeflows)
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(dplyr)
library(ggplot2)
library(tidyr)

``` 


# What should be in the list of reports?
Reports are made on all years available in the DB.
This could be change to concern only the last 5 years (or last 10 years) for which the most data is available.

Report specific:

* Overview report should be based on jfsq level 1

# Create a list of overview reports for the server
This list could itself be a report generated in html form?
Is it possible to create html reports with createreport() ?
Current version of createreport() is rendering in pdf format only:
```
rmarkdown::render(output_format = rmarkdown::pdf_document())
```
Considering that many options in createreport() will not be needed, 
a new function can generate the list of reports as an html file. 
A data frame called "flowavailable" can summarise available information for all possible report types.
A list of all available products by country in all years

|reporter |productcode | period | rawdata | validateddata |
|:--------|:-----------|:-------|:--------|:-------------:|
| France  | 440710     | 2010   | TRUE    | TRUE          |


From this data frame we can extract:

* a list of countries to create links to the overview reports
* a list of products to create links to the completeness reports
* a table of product and country to create links to discrepancy reports

## Summarise flows available
```{r eval=FALSE}
#' Generate a list of distinct reporter and productcode available by year
#' It's rather a lengthy (few seconds) process
#' The output can be used to generate lists of reports
findavailableflows <- function(){
    rawflow <- readdbtbl("raw_flow_yearly") %>%
        select(reporter, productcode, period) %>%
        distinct() %>% 
        collect() %>%
        mutate(rawdata = TRUE)
    validatedflow <- readdbtbl("validated_flow_yearly") %>%
        select(reporter, productcode, period) %>%
        distinct() %>%
        collect() %>%
        mutate(validateddata = TRUE) 
    flowavailable <- full_join(rawflow, validatedflow, by = c("reporter", "productcode", "period"))
    # replace NA value by false
    flowavailable[is.na(flowavailable)] <- FALSE
    return(flowavailable)
}
# system.time(findavailableflows())
#   user  system elapsed 
# 24.308   0.008  42.463 

# Select country only
# system.time(flowavailable <- readdbtbl("validated_flow_yearly") %>%
#         select(reporter) %>%
#         distinct() %>%
#         collect())
#    user  system elapsed 
#   0.024   0.000   7.042 
```


## Generate a list of countries with links to overview reports
```{r eval=FALSE}
# countries <- flowavailable %>% 
#     select(reporter) %>%
#     distinct() %>%
#     mutate(report = paste("<a href='google.com'>search</a>"))
```

A function 


```{r}

```

```{r}
path <- "../../../reports/overview/"

# Simulate country table
countries <- data_frame(reporter = c("Finland", "France")) 
createcountryindex(countries)

# Experiment
# countries %>% 
#     mutate(quantity = paste0("<a href='",path,"overviewquantity",reporter,"jfsq1.pdf'>",
#                                "quantity",reporter,"</a>"),
#            tradevalue = paste0("<a href='",path,"overviewtradevalue",reporter,"jfsq1.pdf'>",
#                                "tradevalue",reporter,"</a>")) %>%
#     kable()

```

