---
title: "Looking at French data series"
output:
  html_document:
    toc: true
---
There was a break in the recording of weight for trade flows reported by 
France between 2006 and 2010. Investigate this issue.


```{r packages, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
library(tradeflows)
library(dplyr)
library(reshape2)
library(ggplot2)
getwd()

``` 




```{r load}
# load raw data and rename it
load("data/sawnwoodexample.rda") # if the format has change,
sawnwood <- sawnwoodexample
sawnwoodexample %>% select(reporter, productcode) %>% unique()
```

## Sawnwood
Load data for france Alexandra said that there is missing weight values between 
2006 and 2010, but I only see missing values for 2012.

Load data from comtrade for an example file of missing data.
Paste this data to Excel.
```{r eval=FALSE}
swdoakfr <- loadcomtrade_bycode(440791, 251, "recent")
unique(swdoakfr$yr)
# add 5 years before
swdoakfr2 <- loadcomtrade_bycode(440791, 251,seq(2004,2008))
swdoakfr <- rbind(swdoakfr, swdoakfr2)
swdoakfr <- swdoakfr %>% arrange(yr, rgCode, ptCode)
unique(swdoakfr$yr[is.na(swdoakfr$NetWeight)])
unique(swdoakfr[c("yr","pfCode")])
xlsx::write.xlsx(swdoakfr,file="data-raw/swdoakfr.xlsx")
# write.xlsx2 is slightly faster on first use. Immensely faster on second use.
system.time(xlsx::write.xlsx(swdoakfr,file="data-raw/swdoakfr.xlsx"))
#    user  system elapsed 
#  14.397   0.056  14.430 
system.time(xlsx::write.xlsx2(swdoakfr,file="data-raw/swdoakfr.xlsx"))
#    user  system elapsed 
#   0.520   0.064   0.601 
```

### Missing data by classification type and country
```{r}
# Data for French sawnwood oak
swdoakfr <- sawnwood %>% 
    filter(productcode == 440791 & reporter=="France") %>%
    select(classification, year, quantity, weight, tradevalue) %>%
    melt(id=c("classification", "year"))  %>%
    mutate(value = !is.na(value)) %>%
    unique

# Add product name and title to the plot
ggplot(data=swdoakfr,aes(x=year, y=variable, 
                         xend=year+1, yend=variable, 
                         color=value, linetype=value)) +
    geom_segment(size=3) + 
    facet_grid(classification~.)
``` 

