---
title: "clean"
output:
  html_document:
    toc : true
---
```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(dplyr)
library(tradeflows)
```

```{r load_data}
load("data-raw/sawnwood_all.RData")
# data frame with comtrade column names
# only for experiments on column renaming
sawnwoodasincomtrade <- swdall
# data frame with efi column names prefere this one for all manipulations
sawnwood <- swdall %>% renamecolumns
```

## Data manipulation
```{r filter}
swd_exp <- filter(sawnwoodasincomtrade, rgDesc == "Export")
swd_imp <- filter(sawnwoodasincomtrade, rgDesc == "Import")
```


## Rename columns
### Using dplyr
I didn't manage to rename columns based on the table
column_names. I used base R to do it, see further down below.
```{r rename_columns}
column_names <- read.csv("data-raw/column_names.csv", as.is=TRUE)
column_names <- column_names %>%
    filter(!is.na(efi) & !is.na(comtrade))
swd <- head(sawnwoodasincomtrade)
# dplyr uses non standard evaluation
swd %>% select(yr, rtCode ) 
# swd %>% select(yr, column_names$comtrade) 
# Error in abs(ind[ind < 0]) : 
#   non-numeric argument to mathematical function
# Standard evaluation
swd %>% select_(.dots = c("yr", "rtCode" )) 
swd %>% select_(.dots = column_names$comtrade) 
# Rename variables
select(swd, year = yr, reporter = rtCode)
select(swd, list(year = yr, reporter=rtCode))
# Issue with quotes
# > select(swd, year = "yr", reporter = rtCode)
# Error in abs(ind[ind < 0]) : 
#   non-numeric argument to mathematical function
# > select(swd, "year = yr, reporter = rtCode")
# Error in abs(ind[ind < 0]) : 
#   non-numeric argument to mathematical function
# > substitute(select(swd, year = y, reporter = rtCode),
substitute(select(swd, year = y, reporter = rtCode),
           list(y = "yr"))
e <- paste("select(sawnwoodasincomtrade, ", "year = yr, reporter = rtCode", ")")
e <- substitute(e)
eval("list(year = yr, reporter=rtCode)")
# sawnwoodasincomtrade %>%
#     rename(product = cmdDescE, reporter = rtTitle ) %>% head
paste(column_names$tradeflows, "=", column_names$comtrade, collapse=", ")
```


### using base R to rename columns 
```{r}
swd <- head(sawnwoodasincomtrade)
names(swd)[names(swd)=="yr"] <- "year"
column_names <- column_names %>%
    filter(!is.na(efi) & !is.na(comtrade))
for (n in column_names["comtrade"][[1]]){
    names(swd)[names(swd)==n] <- 
        column_names$efi[column_names[c("comtrade")]==n]
}
names(swd)

# Try to use a lapply
# swd <- head(sawnwoodasincomtrade)
# rename_column <- function(name, sourcename = "comtrade"){
#     names(swd)[names(swd)==name] <- 
#         column_names$tradeflows[column_names$comtrade==name]
# }
# lapply(column_names["comtrade"][[1]], rename_column)
# names(swd)
``` 

## Calculate unit prices
```{r unitprices}
# sawnwoodasincomtrade <- sawnwoodasincomtrade %>%
#     renamecolumns %>%
#     calcunitprices
# hist(sawnwoodasincomtrade$pricecur)
```

## Add partner trade flows
Check that there are no duplicates
```{r partnerflow}
dtf <- sawnwood 
swap <- dtf %>%
    rename(partnercode = reportercode,
           reportercode = partnercode) %>%
    select(reportercode, partnercode,
           productcode, flow, year,
           quantity, weight, tradevalue) %>%
    mutate(flow = gsub("Import","aaaaaaa",flow),
           flow = gsub("Export","Import",flow),
           flow = gsub("aaaaaaa","Export",flow))
    # Check that values in the dtf and swap tables are all there
    # and in the same order
    stopifnot(dtf$value == swap$value)
    # Merge
    dtf <- merge(dtf, swap, all.x=TRUE, suffixes = c("", "partner"),
                 by = c("reportercode", "partnercode",
                        "productcode", "flow", "year"))
head(dtf)
```

### Check for duplicated flows
There should not be duplicated entires for one combination of
year & flowcode & reportercode & partnercode
```{r duplicates, eval=FALSE}
# Example of a duplicate entry in the swd_all dataset
swdall %>% renamecolumns() %>% select(-productdescription) %>%
    filter(year == 2008 & flowcode ==1 &
                   reportercode == 276& partnercode == 0 )
#   classification year flowcode   flow reportercode reporter reporteriso partnercode
# 1             H3 2008        1 Import          276  Germany         DEU           0
# 2             H3 2008        1 Import          276  Germany         DEU           0
#   partner partneriso productcode unitcode        unit quantity     weight tradevalue
# 1   World        WLD        4407        1 No Quantity       NA 2116672100 1340236000
# 2   World        WLD        4407        1 No Quantity       NA 2116672100 1340236000
print(summary(duplicated(swdall)))
#    Mode   FALSE    TRUE    NA's 
# logical   17185    9291       0 
```


```{r}
# check that there are no duplicates
# swap2 <- unique(swap) # Should be the same length as the rest
swap2 <- swap[duplicated(swap),] %>%
    arrange(year, productcode, flow, partnercode, reportercode)
head(swap2)

swd2 <- dtf[duplicated(dtf),] %>%
    arrange(year, productcode, flow, reportercode, partnercode)
head(swd2) 
# are there duplicates
dtf %>% filter(year == 2008 & flowcode ==1 &
                   reportercode == 276& partnercode == 0 )
```

### Calculate discrepancies
There is one discrepancy column for each column 
weight, quantity and tradevalue.
Calculating it in a molten data frame would facilitate the calculation but
melting a 17000 row data frame returns this 
"Error: C stack usage  9460099 is too close to the limit"
```{r c_stackusageistooclosetolimit, eval=FALSE }
dtf <- swdall %>% renamecolumns %>% removeduplicates %>% addpartnerflow
ids <- names(dtf)[!names(dtf) %in%
                      c("weight", "quantity", "tradevalue",
                        "weightpartner", "quantitypartner",
                        "tradevaluepartner")]
stopifnot(length(ids) + 6 == length(names(dtf)))
dtf2 %>% melt(dtf, id=ids)
# Error: C stack usage  9460099 is too close to the limit
```
