---
title: "Learning to load data from eurostat"
output:
  html_document:
    toc: true
---
Eurostat provides several download facilities, we are looking for 
a bulk download of bilateral trade data.

```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(tradeflows)
library(dplyr)
library(reshape2)
``` 

## Places to download Eurostat data from
### Easy comext
Eurostat [interface to the COMEXT bilateral trade data](http://epp.eurostat.ec.europa.eu/newxtweb/submitformatselect.do) 
enables the selection of reporting and partner area, 
product, trade fow and time period.
For example [Fuel wood Austria](http://epp.eurostat.ec.europa.eu/newxtweb/submitopensavedextraction.do?extractionId=11381457&datasetID=DS-016890&keepsessionkey=true&extractionName=Untitled_20141110&extractionDate=2014/11/10%2014:30:01)


### Bulk download
* Eurostat used to provide [DVD of bulk data](http://epp.eurostat.ec.europa.eu/portal/page/portal/international_trade/publications/comext_dvd). Now available as [bulk download](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&dir=comext%2F2013S2_DVD_image)
* Into the [COMEXT folder](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&dir=comext)
    * [2014S1](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&dir=comext%2F2014S1) contains monthly data since 1988
    * [2014S2](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&dir=comext%2F2014S2) contains yearly data
    * [2014S3](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&dir=comext%2F2014S3) contains montly and yearly data by mode of transport
    * More details in the [Instructions on how to use the bulk download facility.pdf](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&file=comext%2FInstructions+on+how+to+use+the+bulk+download+facility.pdf) and explanation of data format in the [readme.txt](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=3&file=comext%2Freadme.txt)

### SDMX SOAP and Rest APIs
Eurostat provides a [REST API](http://epp.eurostat.ec.europa.eu/portal/page/portal/sdmx_web_services/getting_started/rest_sdmx_2.1#ind_18_1) and a [SOAP API](http://epp.eurostat.ec.europa.eu/portal/page/portal/sdmx_web_services/getting_started/soap_sdmx_2.1).

"These SDMX Web Services are a programmatic access to Eurostat data, with the possibility to: 

* get a complete list of publicly available datasets.
* detail the complete structure definition of a given dataset.
* download a subset of a given dataset or a full dataset. "

## Bulk download of comext data
Yearly data from the [2014S2](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&dir=comext%2F2014S2) folder 
on eurostat's bulk download facility. 


### Download and decompress
Data is downloaded in 7z format. 
It can be uncompressed in Debian GNU linux with the
[p7zip program](https://packages.debian.org/wheezy/p7zip). 
Use system2 to invoque the decompression utility.
Example downloading 2012 data:
```{r, eval=FALSE}
# Download all comext trade flows for a given year
year <- 2012
bulkfile <- paste0("nc",year, "52.7z")
download.file(paste0("http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&file=comext%2F2014S2%2Fdata%2F",bulkfile), 
              destfile = paste0("data-raw/",bulkfile))

# Change to the raw-data directory
curdir <- getwd()  
setwd("data-raw/")

# decompress file and come back to the project root directory
tryCatch(system2("p7zip", args=c("-d", bulkfile)),
    finally = setwd(curdir))
```

### Read .dat file
Comext bulk files are over 400 Mb in size and cannot be read in one slice.
One has to extract a slice of interesting product codes first.

```{r eval=FALSE}
# This takes more than 10 minutes to run on a laptop
# nc2012 <- read.table("data-raw/nc201252.dat", header = TRUE)
bulkfile <- gsub(".7z", ".dat", bulkfile)
productcode <- 4407

# Change to the raw-data directory
curdir <- getwd()  
setwd("data-raw/")
# grep productcode into a smaller file containing only those characters
system2("grep", args=c(productcode,bulkfile), 
        stdout = paste0(productcode,bulkfile))
# Read the first line of bulkfile to use it as col.names 
header <- readLines(bulkfile, n=1)
header <- strsplit(header,",")[[1]]
setwd(curdir)

# read the data file
nc2012 <- read.table("data-raw/4407nc201252.dat", sep=",",
                     col.names = header, as.is=TRUE)
str(nc2012)
# 'data.frame':    24364 obs. of  9 variables:
#  $ DECLARANT    : chr  "001" "001" "001" "001" ...
#  $ PARTNER      : int  3 3 3 3 3 3 3 3 3 3 ...
#  $ PRODUCT_NC   : chr  "44071015" "44071015" "44071031" "44071038" ...
#  $ FLOW         : int  1 2 1 1 1 2 1 2 1 2 ...
#  $ STAT_REGIME  : int  4 4 4 4 4 4 4 4 4 4 ...
#  $ PERIOD       : int  201252 201252 201252 201252 201252 201252 201252 201252 201252 201252 ...
#  $ VALUE_1000ECU: num  64.8 147.4 126.2 213.8 725.1 ...
#  $ QUANTITY_TON : num  44.2 977.3 179.8 207.6 1142.3 ...
#  $ SUP_QUANTITY : int  91 1042 388 498 3891 506 2 311 4218 192 ...
unique(nc2012$PRODUCT_NC) 


```


## Other tools 
* Check the [R sdmx package as mentionned on this question](http://stackoverflow.com/questions/12762431/is-this-the-solution-to-get-data-from-eurostat-into-r)


### Using the SmarterPoland package
Load data from the [Eurostat trade database](http://epp.eurostat.ec.europa.eu/portal/page/portal/international_trade/data/database), using the smarterpoland package

The SmarterPoland package is downloading data tables from the
[bulkdownload listing](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing) but these data tables do not contain bilateral trade.


Example use of the function SmarterPoland::getEurostatRCV
Download the dataset 'Pupil/Student - teacher ratio 
and average class size' from eurostat.
```{r download_educ, eval=FALSE}
tmp <- SmarterPoland::getEurostatRCV(kod = "educ_iste")
head(tmp)
```


Download a sample dataset containing Sawnwood trade by species (for_swspec).
This is downloading from the [Eurostat bulkdownload listing]

* [URL downloading the compressed dataset for_swspec](http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&file=data%2Ffor_swspec.tsv.gz)
```{r download_swd, eval=FALSE}
swd <- SmarterPoland::getEurostatRCV(kod = "for_swspec")
head(swd)
#   treespec prod_wd indic_fo    unit  geo  time      value
# 1    CONIF    SAWN      EXP THS_EUR   AT 2013  1013150.51
# 2    CONIF    SAWN      EXP THS_EUR   BE 2013   229409.43
# 3    CONIF    SAWN      EXP THS_EUR   BG 2013    41869.04
# 4    CONIF    SAWN      EXP THS_EUR BLEU 2013          NA
# 5    CONIF    SAWN      EXP THS_EUR   BR 2013          NA
# 6    CONIF    SAWN      EXP THS_EUR   CA 2013          NA


unique(swd$geo)
#  [1] AT   BE   BG   BLEU BR   CA   CH   CN   CY   CZ   DE   DK   EE   EL   ES   FI  
# [17] FR   HR   HU   ID   IE   IN   IS   IT   LT   LU   LV   ME   MK   MT   NL   NO  
# [33] PL   PT   RO   RU   SE   SI   SK   TR   UK   US   LI  
# 43 Levels: AT BE BG BLEU BR CA CH CN CY CZ DE DK EE EL ES FI FR HR HU ID IE IN ... US


unique(swd[c("treespec", "indic_fo", "unit")]) %>%
    dcast( unit + indic_fo  ~ treespec, value.var="treespec")
#      unit indic_fo C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
# 1 THS_EUR      EXP C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
# 2 THS_EUR      IMP C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
# 3  THS_M3      EXP C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
# 4  THS_M3      IMP C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
# 5 THS_NAC      EXP C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
# 6 THS_NAC      IMP C_FIR CONIF C_PIN NC_ASH NC_BEE NC_BIR NC_CHE NC_MAP NC_OAK NCONIF
#   NC_POP TOTAL
# 1 NC_POP TOTAL
# 2 NC_POP TOTAL
# 3 NC_POP TOTAL
# 4 NC_POP TOTAL
# 5 NC_POP TOTAL
# 6 NC_POP TOTAL
```


Try to download EU trade
as written on this database http://epp.eurostat.ec.europa.eu/portal/page/portal/international_trade/data/database
```{r download_trade, eval=FALSE}
trade <- SmarterPoland::getEurostatRCV(kod = "DS-016890")
head(trade)
```


### Source code of the SmarterPoland package
```{r sourceSmarterPoland, eval=FALSE}
SmarterPoland::getEurostatRCV
# function (kod = "educ_iste") 
# {
#     require(reshape)
#     dat <- getEurostatRaw(kod)
#     dat2 <- t(as.data.frame(strsplit(as.character(dat[, 1]), 
#         split = ",")))
#     cnames <- strsplit(colnames(dat)[1], split = "[,\\\\]")[[1]]
#     colnames(dat2) <- cnames[-length(cnames)]
#     rownames(dat2) <- dat[, 1]
#     rownames(dat) <- dat[, 1]
#     dat3 <- data.frame(dat2, dat[, -1])
#     colnames(dat3) <- c(colnames(dat2), colnames(dat)[-1])
#     dat4 <- melt(dat3, id = cnames[-length(cnames)])
#     colnames(dat4)[ncol(dat4) - 1] = cnames[length(cnames)]
#     dat4
# }
# <environment: namespace:SmarterPoland>

SmarterPoland::getEurostatRaw
# function (kod = "educ_iste") 
# {
#     adres <- paste("http://epp.eurostat.ec.europa.eu/NavTree_prod/everybody/BulkDownloadListing?sort=1&file=data%2F", 
#         kod, ".tsv.gz", sep = "")
#     tfile <- tempfile()
#     download.file(adres, tfile)
#     dat <- read.table(gzfile(tfile), sep = "\t", na.strings = ": ", 
#         header = F, stringsAsFactors = F)
#     unlink(tfile)
#     colnames(dat) <- as.character(dat[1, ])
#     dat <- dat[-1, ]
#     for (i in 2:ncol(dat)) {
#         tmp <- sapply(strsplit(as.character(dat[, i]), split = " "), 
#             `[`, 1)
#         tmp[tmp == ":"] = NA
#         dat[, i] <- as.numeric(tmp)
#     }
#     dat
# }
# <environment: namespace:SmarterPoland>
``` 

