---
title: "Explore ggplot2 functionalities"
author: "Paul Rougieux"
date: "27/11/2014"
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true
---


```{r packages, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
```
Explore some of the ggplot2 functions for use in the project


## Sample plot
```{r irisplot}
ggplot(iris, 
       aes(x=Petal.Length, y=Petal.Width, color=Species)) + 
    geom_point()
```

## Alpha

### Alpha to avoid overplotting
Use aesthetic to make points more transparent on a coutry * contry grid 
display trade volume as alpha level.
```{r alphaoverplot}
df <- data.frame(x = rnorm(5000), y = rnorm(5000))
ggplot(df, aes(x,y)) + geom_point(alpha = 1/10)
```

### Alpha to add shading
```{r alphshading}
b <- ggplot(economics, aes(x = date, y = unemploy))
j <- b + geom_line()
# j # uncomment to display basic plot
yrng <- range(economics$unemploy)
j <- j + geom_rect(aes(NULL, NULL, xmin = start, xmax = end, fill = party),
ymin = yrng[1], ymax = yrng[2], data = presidential)
# j
library(scales) # to access the alpha function
j + scale_fill_manual(values = alpha(c("blue", "red"), .3))
```

### Scales for color and linetype
For some countries, the completeness table contained only true values. 
WHIch were rendered as red lines, for other countries 
true values were rendered as blue lines. 
Changed this behaviour by using 
manual color palettes for `colours scale_colour_manual()`
and line palette for lines `scale_linetype_manual()`.

```{r scalesggplot}
# plot completeness
rawdata <- tradeflows::sawnwoodexample
r <- "France"
productcodeinreport <- 4407
completeness <- rawdata %>%
    filter(productcode == 4407) %>%
    select(reporter, classification, year, quantity, weight, tradevalue) %>%
    melt(id=c("reporter", "classification", "year"))  %>%
    mutate(value = !is.na(value)) %>%
    unique %>% 
    arrange(classification, reporter, year)

    p <- ggplot(data=filter(completeness, reporter == r), 
                aes(x=year, y=variable, 
                    xend=year+1, yend=variable, 
                    color=value, linetype=value)) +
        scale_colour_manual(values = c("TRUE" = "blue", "FALSE" = "red")) +
        scale_linetype_manual(values = c("TRUE" = 2, "FALSE" = 1)) +
    #"solid", "dashed"
        geom_segment(size=3) + 
        facet_grid(classification~.) +
        ggtitle(paste("Data completeness of", productcodeinreport,
                      "in", r))
    plot(p)

```

## Shapes
```{r shapes}

ggplot(data.frame(x = seq(1,6)), aes(x = x,y =x,shape=as.factor(x),color=as.factor(x))) + geom_point(size=10)

```


## Two plots with grid.arrange() sharing one legend
Seen in [Hadley's wiki](https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs)
and in this [Stackoverflow question](http://stackoverflow.com/questions/13649473/add-a-common-legend-for-combined-ggplots/28594060#28594060)
```{r}
library(ggplot2)
library(grid) # to use unit.c
library(gridExtra)

grid_arrange_shared_legend <- function(...) {
    plots <- list(...)
    g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
    legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    grid.arrange(
        do.call(arrangeGrob, lapply(plots, function(x)
            x + theme(legend.position="none"))),
        legend,
        ncol = 1,
        heights = unit.c(unit(1, "npc") - lheight, lheight))
}

dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
p1 <- qplot(carat, price, data=dsamp, colour=clarity)
p2 <- qplot(cut, price, data=dsamp, colour=clarity)
p3 <- qplot(color, price, data=dsamp, colour=clarity)
p4 <- qplot(depth, price, data=dsamp, colour=clarity)
grid_arrange_shared_legend(p1, p2, p3, p4)
     
```
