---
title: "Development scrap concerning the input data"
output:
  html_document:
    toc: true
---


```{r packages}
require(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
require(tradeflows)
require(RJSONIO) # Why are the plyr and dplyr packages not loaded whith treadeflows?
# This didn't work before because you need to properly document the function in roxygen
# placing @export before the function creation instruction 
# then create a NAMESPACE file with devtools::document()
require(plyr)
# require(dplyr)
# Load plyr first to avoid this message:
# -------------------------------------------------------------------------------
#     You have loaded plyr after dplyr - this is likely to cause problems.
# If you need functions from both plyr and dplyr, please load plyr first, then dplyr:
#     library(plyr); library(dplyr)
# -------------------------------------------------------------------------------
```


## UN comtrade API parameters
The [UN Comtrade API](http://comtrade.un.org/data/doc/api/) has a limit of 50 000 records per request. But the max parameter is set by default to get only 500 records see for example, the time period below.


### ps Time Period
To download many years at once, paste them in a character string, separated by commas.
The max parameter defaults to 500. 
Illustration of a problem encountered as I had forgotten to set the max limit.
```{r time_period, eval=FALSE}
swd89101112 <- tradeflows::loadcomtrade_bycode(4407, 276, "2012,2011,2010,2009,2008", max=500)
# Year 2010 and 2011 are not in this table.
unique(swd89101112$yr)
# [[1]]
# [1] 2012
# 
# [[2]]
# [1] 2008
# 
# [[3]]
# [1] 2009


# But 2010 and 2011 are available in the comtrade API.
swd10_11 <- loadcomtrade_bycode(4407, 276, "2011,2010")
unique(swd10_11$yr)
# trying URL 'http://comtrade.un.org/api/get?cc=4407&r=276&p=all&rg=all&ps=2011,2010&px=HS&max=50000&fmt=json'
# Content type 'application/json; charset=utf-8' length 177788 bytes (173 Kb)
# opened URL
# ==================================================
# downloaded 173 Kb
# 
# > unique(swd10_11$yr)
# [[1]]
# [1] 2010
# 
# [[2]]
# [1] 2011
```


### px Classification
```{r classification}

```


### r Reporting Area
Try to load for all countries
```{r reporter, eval=FALSE}
swd <- loadcomtrade_bycode(4407, "all", 2013)

# trying URL 'http://comtrade.un.org/api/get?cc=4407&r=all&p=all&rg=all&ps=2013&px=HS&max=50000&fmt=json'
# 
# Content type 'application/json; charset=utf-8' length 405 bytes
# opened URL
# ==================================================
# downloaded 405 bytes
# 
# Quitting from lines 50-51 (input.Rmd) 
# Error in loadcomtrade_bycode(4407, "all", 2013) : 
#   list(name = "Query complexity", value = 5002, category = 0, description = "your query is too complex. Please simplify.", helpUrl = "")Both 'all' reporters and 'all' partners may not be selected. Select a different reporter or partner.list(value = 0, started = "0001-01-01T00:00:00", finished = "0001-01-01T00:00:00", durationSeconds = 0)NA
``` 


## Converting JSON data to a data.frame
```{r load_json_file}
json <- fromJSON("data-raw/UNCOMTRADE_exp.JSON", nullValue = NA)
stopifnot(json$validation$status$name == "Ok") # Check status
```


### Convert to a data.frame using do.call
```{r convert_do_call}
# head() keeps only the first few lines
swd <- as.data.frame(do.call("rbind", head(json$dataset)))
# When I use do.call, the
# structure of the data frame looks bizarre, more like a list than a dataframe.
str(swd)
# Structure of this binded list created by do.call
str(do.call("rbind", head(json$dataset)))
```


### Convert to a data frame using a loop
```{r convert_loop}
# Use a loop to create data frame first then bind them together
swd2 <- data.frame()
system.time(
    for (obs in json$dataset){
        swd2 <- rbind(swd2,data.frame(obs))
        })
str(swd2)
```


### Convert to a data.frame using ldply
```{r convert_ldply}
system.time(swd3 <- ldply(json$dataset, function(l) data.frame(l)))
str(swd3)
```


## Manipulate Comtrade data
```{r filter}
swd_exp <- filter(swd, rgDesc == "Export")
swd_imp <- filter(swd, rgDesc == "Import")
``` 
