---
title: "ITTO product classification"
output:
  html_document:
    toc: true
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
# opts_chunk$set(echo=FALSE)
library(tradeflows)
library(dplyr)
``` 


## Load the ITTO product table and rename columns
ITTO uses internal product names.
The classification is different to what was used in comtrade.
_HS07_ in this ITTO table probably corresponds to _HS3_ in comtrade
that is why we let it called nomenclature for the moment.

Description is called product in the efi convention for column names,
this might have to be changed.

```{r readcsvandrename}
itto <- read.csv("data-raw/ITTO_products.csv", as.is=TRUE)
str(itto)
itto <- itto %>%
    select(product = Names.of.products, 
           productcodeitto = Product.code, 
           nomenclature = Nomenclature,
           productcodecomtrade = Code,
           description = Description)
```

## List ITTO product codes and names
```{r rename}
itto %>% 
    select(product, productcodeitto) %>%
    unique() %>%
    kable()
```

## Compare ITTO code and names to those of Comtrade
### Example products under ITTO SAWNWOOD
```{r sawnwood}
swd <- itto %>% 
    filter(product == "SAWNWOOD" )
kable(swd)
```

### Example products under Comtrade 4407
4407 is an aggregate of sawnwood products and doesn't appear
in the ITTO table
```{r sawnwood4407}
itto %>%
    filter(productcodecomtrade == 4407) 
```

### Example products under Comtrade 440792
For ITTO, 440792 is a sub product of sawnwood.
```{r sawnwoodbeech}
itto %>%
    filter(productcodecomtrade == 440792) %>%
    kable()
```

### Example products under Comtrade 440799
```{r sawnwoodbeech}
itto %>%
    filter(productcodecomtrade == 440799) %>%
    kable()
```
