---
title: "Learning to parse json files from comtrade"
output:
  html_document:
    toc: true
---

 
This document is relating development steps towards converting 
JSON data from the UN comtrade database to R data frames. 

I have been inspired by several online reading, from package authors 
and by questions on stackoverflow:
[Converting a list of data frames into one data frame in R](http://stackoverflow.com/questions/2851327/converting-a-list-of-data-frames-into-one-data-frame-in-r).  After unsatisfactory trials with the rjson and RJSONIO packages. The most crucial step was discovering the jsonlite package. jsonlite transforms a json file directly into a list of data frames.

```{r packages}
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(tradeflows)
library(RJSONIO)
library(jsonlite)

# Why are the plyr and dplyr packages not loaded whith treadeflows?
# This didn't work before because you need to properly document the function in roxygen
# placing @export before the function creation instruction 
# then create a NAMESPACE file with devtools::document()
# Will make the required package autoload.
library(plyr)
library(dplyr)
# Load plyr before dplyr to avoid this message:
# -------------------------------------------------------------------------------
#     You have loaded plyr after dplyr - this is likely to cause problems.
# If you need functions from both plyr and dplyr, please load plyr first, then dplyr:
#     library(plyr); library(dplyr)
# -------------------------------------------------------------------------------
```


## Converting JSON data to a data.frame using RJSONIO::fromJSON
```{r load_json_file}
warning("If data-raw/UNCOMTRADE_exp.JSON file missing, look at the unevaluated peace of code above to download the JSON file.")
json <- RJSONIO::fromJSON("data-raw/UNCOMTRADE_exp.JSON", nullValue = NA)
stopifnot(json$validation$status$name == "Ok") # Check status
```


### Convert the list to a data.frame using do.call
```{r convert_do_call}
# head() keeps only the first few lines
swd <- as.data.frame(do.call("rbind", head(json$dataset)))
# When I use do.call, the
# structure of the data frame looks bizarre, more like a list than a dataframe.
str(swd)
# Structure of this binded list created by do.call
str(do.call("rbind", head(json$dataset)))
```


### Convert the list to a data frame using a loop
```{r convert_loop}
# Use a loop to create data frame first then bind them together
swd2 <- data.frame()
system.time(
    for (obs in json$dataset){
        swd2 <- rbind(swd2,data.frame(obs))
        })
str(swd2)
```


### Convert the list to a data.frame using ldply
This gives a nice table structure
```{r convert_ldply}
system.time(swd3 <- ldply(json$dataset, function(l) data.frame(l)))
str(swd3)
```


### Convert the list to a data.frame using sapply
The structure is still not nice.
```{r}
system.time(swd4 <- data.frame(t(sapply(head(json$dataset), data.frame))))
str(swd4)
```


### Converting JSON data to a data.frame using jsonlite::fromJSON
Gives a nice data.frame structure
```{r}
system.time(json2 <- fromJSON("data-raw/UNCOMTRADE_exp.JSON"))
stopifnot(json2$validation$status$name=="Ok")
swd5 <- json2$dataset
str(swd5)
```

