---
title: "Learning to read and write data into SQLite"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tradeflows)
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(dplyr)
``` 


# Convert from MySQL to SQLite 

[conversion scirpt](https://gist.github.com/esperlu/943776#file-mysql2sqlite-sh) found on a [Stackoverflow answer](http://stackoverflow.com/a/9933603/2641825).
See also below for another version of that script. 

## Conversion script first version 
The script connects to MySQL, creates the dump 
and then convers to SQLite.
```
cp ~/downloads/mysql2sqlite.sh /tmp/
cd /tmp
chmod +x mysql2sqlite.sh # Make it executable
# Old version by Esperlu
./mysql2sqlite.sh -u R -p tradeflows | sqlite3 tradeflows.sqlite
```
The raw database was copied but there was an issue with the validated database: `Error: near line 5217030: no such table: validated_flow_yearly`.

## Conversion script second version 
Another version of the conversion script is 
maintained at https://github.com/dumblob/mysql2sqlite
It converts from a MySQL dump into a SQLite dump.
```
# Same as above for download and permission
# Command is different
cd /tmp
chmod +x mysql2sqlite.sh
./mysql2sqlite.sh raw_ | sqlite3 tradeflows.sqlite
```

## Importing from CSV into SQLite
[How to output MySQl query result in CSV format ?](http://stackoverflow.com/questions/356578/how-to-output-mysql-query-results-in-csv-format)

MySQL command to export CSV
```
select * from raw_flow_yearly into outfile '/tmp/raw_flow_yearly2.csv' fields terminated by ',' enclosed by '"' lines terminated by '\n';
```
Copy database structure.
[SQLite datatypes](https://www.sqlite.org/datatype3.html) are called affinity.
Since the csv file doesn't contain information 
about the datatypes, I decided to copy the structure from MYSQL into SQLite and use that structure. 



SQLite command to import CSV



## Browse the newly created SQLite database 
```
# Bash command to start sqlite
sqlite3 tradeflows.sqlite3
# Commands within sqlite
.tables
.schema raw_flow_yearly
select * from raw_flow_yearly limit 5;

```

# Using dplyr to read from SQLite

```{r}
tf_sqlite <- src_sqlite("/tmp/tradeflows.sqlite3")
tf <- tbl(tf_sqlite, "raw_flow_yearly")

system.time(repprod <- tf %>% select(reporter, productcode, period) %>%
    distinct() %>% collect())
#  user  system elapsed 
#   3.908   0.660   4.566 
```


# Compare with the same operation on MySQL
```{r}
tf2 <- readdbtbl("raw_flow_yearly")
system.time(repprod2 <- tf2  %>% 
                select(reporter, productcode, period) %>%
                distinct() %>% collect())
#  user  system elapsed 
#   0.072   0.004  13.408 

# Compare the output
repprod <- repprod %>% arrange(reporter, productcode, period)
repprod2 <- repprod2 %>% arrange(reporter, productcode, period)
summary(repprod$productcode - repprod2$productcode)
```


# Using dplyr to write into SQLite 
```{r}
tfdb <- readdbtbl("validated_flow_yearly")
swd99 <- readdbproduct(440799, "validated_flow_yearly")
tfdblite <- src_sqlite("/tmp/tradeflows.sqlite3", create = T)


tf_lite <- copy_to(tfdblite, swd99, temporary = FALSE, 
                   indexes = list("year", "flowcode","reportercode","partnercode","productcode"))
```

