---
title: "Montly Unit Prices"
author: "Paul Rougieux"
date: "04/12/2014"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE, warning=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
library(lubridate)
``` 

```{r loadandprepare}
prepare <- function(dtf){
    dtf %>% 
        # Keep only import and export
        filter(flowcode %in% c(1,2)) %>%
        # Remove EU28 reporter
        filter(!reporter %in%c("EU-27")) %>%
        # Remove World partner
        filter(!partner %in%c("EU-27", "World")) %>%
        merge(select(reportercomtrade, reportercode,
                     regionreporter=region)) %>%
        merge(select(reportercomtrade, 
                     partnercode = reportercode,
                     regionpartner=region)) %>%
        mutate(period = parse_date_time(period, "%y%m"),
               price = tradevalue / quantity,
               pricew = tradevalue / weight,
               # on sum of all values
               tradevalue = as.numeric(tradevalue),
               # To avoid Error in swd99$c(NA_integer_,  : 
               # invalid subscript type 'integer'
               quantity = as.numeric(quantity),
               # tradeflow within a region 
               withinregion =  regionreporter == regionpartner)
    }

swd99 <- readdbproduct(440799, "raw_flow_monthly")  %>% 
    prepare
```

```{r eval=TRUE}
# Load data for another product
swd <- classificationcomtrade$HS %>% filter(parentcode == "4407")

# Load oak data from comtrade, to a file
# comment out to run this (only once, lasts 10 minutes)
# loadcomtradewithpause(440791, path = "data-raw/", freq = "M")
# load file
load("data-raw/440791.RData")
oak <- dtf %>% renamecolumns %>% prepare

# Here we have the quantity
summary(oak[c("tradevalue", "quantity", "weight",
              "price", "pricew")])
unique(oak$productdescription)
```

### Monthly prices in European countries
```{r}
swd99eu <- swd99 %>% 
    filter(regionreporter == "Europe") %>%
    group_by(flow, reporter, period) %>% 
    summarise(medianprice = median(price, na.rm=TRUE),
              weight = sum(weight, na.rm=TRUE),
              tradevalue = sum(tradevalue)) %>%
    mutate(price = weight/tradevalue, 
           tradevaluecut = round(log10(tradevalue)))

swd99 <- swd99 %>% 
    mutate(tradevaluecut = round(log10(tradevalue)))
summary(swd99$price)
summary(log10(swd99$tradevalue))

# Point cloud, all prices
p <- ggplot(filter(swd99, regionreporter == "Europe"),
       aes(x = period, y = price, color = reporter)) +
    ylim(0,6) + geom_point(alpha = 1/10) +
    ggtitle("Monthly prices for all 
            available trade flows in Europe")
p + facet_grid(~flow)
p + facet_grid(tradevaluecut~flow)    

# Point cloud, average prices for one country
p <- ggplot(swd99eu,
            aes(x = period, y = price, color = reporter)) +
    ylim(0,6) + geom_point(alpha = 1/10) +
 
    ggtitle("Average prices per reporting country")
p + facet_grid(~flow) 
p + facet_grid(tradevaluecut~flow)

# Distribution of import prices along the trade volume
p <- ggplot(filter(swd99, reporter %in% largeimp),
       aes(x = period, y = price, color = reporter)) +
    ylim(0,6) + geom_point(alpha = 1/10) +
    ggtitle("Monthly prices for all 
            available trade flows in Europe")

```


### Import prices for large importers
```{r}
largeimp <- c("Italy", "Belgium", "Germany", "United Kingdom",
              "Austria", "Netherlands", "France", "Denmark",
              "Portugal", "Poland")
p <- ggplot(filter(swd99, reporter %in% largeimp),
       aes(x = period, y = price, color = reporter)) +
    ylim(0,6) + geom_point(alpha = 1/10) +
    ggtitle("Monthly prices for all 
            available trade flows in Europe")
p + facet_grid(reporter~flow) +
    geom_line(data = filter(swd99eu, reporter %in% largeimp))
```

### Oak import prices for large importers
```{r}
p <- ggplot(filter(oak, reporter %in% largeimp),
       aes(x = period, y = price, color = reporter)) +
    ylim(0,2000) + geom_point(alpha = 1/10) +
    ggtitle("Monthly prices for all 
            available trade flows in Europe")
p + facet_grid(reporter~flow) 
```

## Exploration of the dataset
There are no quantities, therefore we cannot calculate cubic meter prices. But weight and trade values are available, prices will be in US $ per kilogram. 
```{r}
summary(swd99[c("quantity", "weight", "tradevalue")])
```


### Largest exporters and importers
Largest exporters and importers over the period
```{r}
large <- swd99 %>% filter(regionreporter == "Europe") %>%
    group_by(flow, reporter) %>%
    summarise(tradevalue = sum(tradevalue),
              weight = sum(weight)) %>% 
    data.frame %>%
    arrange(desc(tradevalue))
largeexp <- large %>% filter(flow == "Exports") %>% head(10) 
largeexp %>% kable
largeimp <- large %>% filter(flow == "Imports") %>% head(10)
largeimp %>% kable
``` 


