---
title: "Price distribution"
author: "Paul Rougieux"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
---

```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE, warning=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 

Looking at softwood lumber prices in Europe, after reading a publication by Wei-Yew Chang and Chris Gaston, [The competitiveness of Canadian softwood lumber a disaggregated trade-flow analysis](http://www.nrcresearchpress.com/doi/abs/10.1139/cjfr-2014-0058).
In this publication they chose a limit of 148US$/m-3 between low
and high quality lumber.

```{r loadandprepare}
prepare <- function(dtf){
    dtf %>%
        filterworldeu28() %>%
        addregion() %>%
        addconversionfactorandprice() %>%
        mutate(withinregion =  regionreporter == regionpartner)    
}
swd10 <- readdbproduct(440710, "raw_flow_yearly") %>% prepare
swd91 <- readdbproduct(440791, "raw_flow_yearly") %>% prepare
```

# Price distribution by region
## Price distribution by quantity
```{r}
pricepoint <- ggplot(swd10, aes(x = price, y = log(quantity))) + 
    geom_point() + xlim(0,1000) 
pricepoint + facet_grid(regionpartner ~ flow) +
    aes(color=regionreporter)
pricepoint + facet_grid(regionpartner ~ regionreporter) +
    aes(color = flow)

swd10region <- swd10 %>%
    group_by(regionreporter, regionpartner, year, flow) %>%
    summarise(quantity = sum(quantity, na.rm=TRUE),
              tradevalue = sum(tradevalue, na.rm = TRUE),
              price = tradevalue / quantity)
ggplot(swd10region, aes(x = price, y = log(quantity),
                        color = flow)) +
    geom_point() + xlim(0,1000) +
    facet_grid(regionpartner ~ regionreporter) 
```

### European import prices
For a selection of countries
```{r}

countrylist1 <-c("France","Germany","Sweden", "United Kingdom")
countrylist2 <- c("Germany", "Austria", "Finland", "Sweden",
                  "Estonia", "Latvia", "Lithuania")    
swd10eu <- swd10 %>%
    filter(reporter %in% countrylist2 &
               partner %in% countrylist2) %>%
    select(reporter, partner, flow,
           year, price, quantity, tradevalue) %>%
    arrange(price)
write.csv(swd10eu, file = "/tmp/sawnwood10_frdeswuk.csv")

pricepoint <- ggplot(swd10eu, aes(x = price, y = log(quantity))) + 
    geom_point() + xlim(0,1000) 
pricepoint + facet_grid(partner ~ reporter) +
    aes(color = flow)
```


## Oak price distribution
```{r}
pricepoint <- ggplot(swd91, aes(x = price, y = log(quantity))) + 
    geom_point() + xlim(0,3000) 
pricepoint + facet_grid(regionpartner ~ regionreporter) +
    aes(color = flow)
```


