---
title: "Unit Prices"
author: "Paul Rougieux"
date: "04/12/2014"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE, warning=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 

```{r loadandprepare}
# Remove this chunk in the template file
# load("data-raw/comtrade/440799.RData")
# replace file by database source
dtf <- readdbproduct(440799, "raw_flow_yearly")
regionpartner <- reportercomtrade %>%
    select(reportercode = reportercode, regionpartner=region)
swd99 <- dtf %>% 
#     renamecolumns %>%
    filter(flow %in% c("Import", "Export")) %>%
    # Remove EU28 reporter
    filter(!reporter %in%c("EU-28")) %>%
    # Remove World partner
    filter(!partner %in%c("EU-28", "World")) %>%
    # keep region 
    # but as of 23 December 2014, use regionreport and regionpartner instead
    merge(select(reportercomtrade, reportercode, region, subregion)) %>%
    merge(select(reportercomtrade, reportercode, regionreporter=region)) %>%
    merge(select(reportercomtrade, 
                 partnercode = reportercode, regionpartner=region)) %>%
    mutate(conversion = weight / quantity,
           price = tradevalue / quantity,
           # To avoid "integer overflow - use sum(as.numeric(.))" error 
           # on sum of all values
           tradevalue = as.numeric(tradevalue),
           # To avoid Error in swd99$c(NA_integer_,  : 
           # invalid subscript type 'integer'
           quantity = as.numeric(quantity),
           # tradeflow within a region 
           withinregion =  regionreporter == regionpartner)
```



## Number of flows and total trade value of missing quantity and weight data
### Flows for which there is no quantity
Not all quantities are available. which percentage of the trade value cannot be taken into account for the unit price calculation? 

* There are `r nrow(swd99)` flows in the  `r as.character(unique(swd99$productcode))` dataset. 
* `r nrow(filter(swd99, is.na(tradevalue)))` don't have a trade value
* `r nrow(filter(swd99, is.na(weight)))` don't have a weight
* `r nrow(filter(swd99, is.na(quantity)))` don't have a quantity

```{r percentagenoquantity}
swd99 %>% 
    mutate(has_a_quantity = !is.na(quantity)) %>%
    group_by(flow, has_a_quantity) %>%
    summarise(number_of_flows = n(),
              total_tradevalue = sum(tradevalue, na.rm=TRUE)) %>%
    mutate(percentage_of_value = 
               round(total_tradevalue / sum(total_tradevalue, na.rm=TRUE)*100)) %>%
    arrange(-has_a_quantity) %>%
    kable
```

### Flows for which there is no weight
```{r}
swd99 %>% 
    mutate(has_a_weight = !is.na(weight)) %>%
    group_by(flow, has_a_weight) %>%
    summarise(number_of_flows = n(),
              total_tradevalue = sum(tradevalue, na.rm=TRUE)) %>%
    mutate(percentage_of_value = 
               round(total_tradevalue / sum(total_tradevalue, na.rm=TRUE)*100)) %>%
    arrange(-has_a_weight) %>%
    kable
```



## Unit prices at various aggregation level
Calculate prices for product code `r as.character(unique(swd99$productcode))`.
Calculate median prices and an upper and lower bound. Upper price is the third quartile multiplied by 2, lower price is the first quartile divided by 2.
In a boxplot, the upper and lower "hinges" correspond to the first and third quartiles (the 25th and 75th percentiles).

### Median unit prices by regions
```{r medianpriceregion, results='asis', fig.height=6}
swd99region <- swd99 %>%
    group_by(flow, region, year) %>%
    summarise(lowerprice = round(0.5 * quantile(price, 0.25, 
                                                names=FALSE, na.rm=TRUE)),
              medianprice = round(median(price, na.rm=TRUE)),
              upperprice = round(2 * quantile(price, 0.75,
                                              names=FALSE, na.rm=TRUE))) %>%
    arrange(-medianprice)

swd99region %>%
    dcast(region + flow  ~ year, value.var="medianprice") %>%
    kable(caption="Median unit prices")


ggplot(data=swd99region, aes(x=year, y=medianprice, color=flow)) +
    coord_cartesian(ylim=c(0,2500)) +
    geom_ribbon(aes(ymin=lowerprice, ymax=upperprice, 
                    fill=flow, alpha=0.05))+
    geom_point(colour="grey50", size = 4) + geom_point(size=3) +
    ggtitle("Median prices by region, shaded area represents 
            a zone between our limits on lowerprice and upperprice") +
    facet_wrap(region~flow) 
cat("\n\n")
p <- ggplot(data=swd99, aes(x=factor(year), y=price)) +
    geom_boxplot(na.rm=TRUE) + xlab("") +
    facet_wrap(region~flow)  
p + coord_cartesian(ylim=c(0,2500)) + ggtitle( "440799 zoom plot (price<2500$)") 
cat("\n\n")
p + scale_y_log10() + ggtitle( "440799 Log scale") 


# Do we need a separate calculation because of NA quantity
# compare to Unit price of aggregated trade flows
swd99regionunitprice <- swd99 %>%
    filter(!is.na(quantity)) %>%
    group_by(region, flow, year) %>%
    summarise(price = round(sum(tradevalue) / sum(quantity))) 
swd99regionunitprice %>%
    dcast(region + flow~year, value.var="price") %>%
    kable(caption="Unit prices of aggregated trade flows")
```


### Median unit price within a region and region with world
```{r medianpriceswithinregion}
# Median prices
swd99region2 <- swd99 %>%
    group_by(withinregion, flow, region, year, flag) %>%
    summarise(lowerprice = round(0.5 * quantile(price, 0.25, 
                                                names=FALSE, na.rm=TRUE)),
              medianprice = round(median(price, na.rm=TRUE)),
              upperprice = round(2 * quantile(price, 0.75,
                                              names=FALSE, na.rm=TRUE))) %>%
    arrange(-medianprice)

ggplot(swd99region2,
       aes(x=year, y=medianprice, color=withinregion)) +
    geom_point() + facet_grid(flow+flag~ region) + 
    ylim(c(0,1000))



swd99region2 %>%
    dcast(withinregion + region + flow  ~ year, value.var="medianprice") %>%
    kable(caption="Median unit prices")
```


### Median unit price and comtrade estimated quantities
```{r}
# medianprice

```


### Median unit prices by subregions
```{r medianpricesubregion, results='asis', fig.height=6}
# Table of median flow prices
swd99subregion <- swd99 %>% 
    group_by(flow,subregion, year) %>%
    summarise(lowerprice = round(0.5 * quantile(price, 0.25, 
                                                names=FALSE, na.rm=TRUE)),
              medianprice = round(median(price, na.rm=TRUE)),
              upperprice = round(2 * quantile(price, 0.75,
                                              names=FALSE, na.rm=TRUE))) %>%
    arrange(-medianprice)
kable(swd99subregion)

```


#### Analyse upper and lower prices by subregions in 2011.

```{r medianpricessubregion2011}
# Transform subregion to an ordered vector by import prices
f = "Import"
swd99imp <- swd99subregion %>%
    as.data.frame %>% # Otherwise grouping variables are used before arrange() 
    filter(flow==f &year==2011) %>% arrange(-medianprice)


# Add these lower, upper and median prices on the plot
swd99subregionmelt <- swd99subregion %>% 
    as.data.frame %>% #Prevent message cannot modify grouping variable
    filter(year==2011) %>%
    mutate(subregion = ordered(subregion, levels = swd99imp$subregion)) %>%
    melt(id=c("flow", "subregion", "year")) %>%
    mutate(variable = ordered(variable, levels=c("upperprice",
                                                 "medianprice","lowerprice" )))
ggplot(data=swd99subregionmelt) +
    geom_point(aes(x = subregion, y = value, color =  variable))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_grid(flow~.)


p <- ggplot(data=swd99, aes(x=subregion, y=price)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    geom_point(data=swd99subregionmelt,
             aes(x = subregion, y = value, color =  variable))+
   facet_grid(flow~.) + xlab("")
cat("\n\n")
p + coord_cartesian(ylim=c(0,2500)) + ggtitle( "440799 zoom plot (price<2500$)") 

cat("\n\n")
p + scale_y_log10() + ggtitle( "440799 Log scale") 
```

## Distributions
```{r}
summary(select(swd99, tradevalue, weight, quantity, price))
```

### Distribution of trade value
Slice the dataset in log and show the distribution in number of rows and in 
value along the trade value split in log.
```{r}
round(log(max(swd99$tradevalue)))
swd99logtv <- swd99 %>% 
    mutate(logtradevalue = round(log(tradevalue))) %>%
    group_by(logtradevalue, region, withinregion) %>%
    summarise(tradevalue = sum(tradevalue), number_of_flows = n()) %>%
    melt(id=c("logtradevalue", "region", "withinregion")) 
p <- ggplot(swd99logtv, aes(x=logtradevalue, y=value)) + 
    geom_bar(stat="identity") + ylab("Number of flows | Trade value in $")
p + facet_grid(variable~., scales="free_y")
p + facet_grid(variable~region, scales="free_y") 
p + facet_grid(variable+withinregion~region, scales="free_y") +
    ggtitle("Trade within the region (TRUE) 
            or with a partner outside the region (FALSE)")
```

### Distribution of quantity 
```{r}
round(log(max(swd99$quantity)))
swd99logq <- swd99 %>% 
    mutate(logquantity = round(log(quantity))) %>%
    group_by(logquantity, region) %>%
    summarise(quantity = sum(quantity), number_of_flows = n()) %>%
    melt(id=c("logquantity", "region")) 

p <- ggplot(filter(swd99logq, !is.na(logquantity)&logquantity!=-Inf), 
            aes(x=logquantity, y=value)) + 
    geom_bar(stat="identity") + ylab("Number of flows | Quantity in m3")
p + facet_grid(variable~., scales="free_y")
p + facet_grid(variable~region, scales="free_y") 
```

### Distribution of weight 
```{r}
swd99logw <- swd99 %>% 
    mutate(logweight = round(log(weight))) %>%
    group_by(logweight, region) %>%
    summarise(weight = sum(weight), number_of_flows = n()) %>%
    melt(id=c("logweight", "region")) 

p <- ggplot(filter(swd99logw, !is.na(logweight)&logweight!=-Inf), 
            aes(x=logweight, y=value)) + 
    geom_bar(stat="identity") + ylab("Number of flows | weight in kg")
p + facet_grid(variable~., scales="free_y")
p + facet_grid(variable~region, scales="free_y") 
```



### Distribution of unit prices between `r min(swd99$year)` and `r max(swd99$year)`
```{r pricedistributionnrow}
p <- ggplot(swd99, aes(x=price)) + 
    geom_histogram(binwidth=100) + xlim(0,2500) +
    ylab("Number of trade flows")
p + facet_grid(flow~.) + ggtitle("All world flows for 440799")
p + facet_grid(flow~withinregion+flag) + ggtitle("All world flows for 440799")
# p + facet_grid(region~flow) + ggtitle("World flows for 440799 by regions")
# p + facet_grid(year~flow) + ggtitle("World flows for 440799 by year")
p + facet_grid(flow + year ~flag) + ggtitle("World flows for 440799 by flag")
p + facet_grid(region + flow ~ year) +
    ggtitle("Number of flows for 440799 by region and year
            vertical red line represents median price") + 
    geom_vline(data=swd99region, aes(xintercept = medianprice), color="red")
```


```{r pricedistributiontradevalue, fig.height=10}
# Calculate total trade value by region, year and show total trade value
p <- swd99 %>% mutate(pricegroup = round(price/100) * 100) %>%
    group_by(region, flow, year, pricegroup) %>% 
    summarise(tradevalue = sum(tradevalue)) %>%
    filter(!is.na(pricegroup)&pricegroup!=Inf) %>%
    ggplot(aes(x=pricegroup,y=tradevalue)) + geom_bar(stat="identity") +
    ylab("Total trade value in $") +
    xlim(0,2500) + xlab("Price in $") 
cat("\n\n")
p + facet_grid(flow~.) + ggtitle("All world flows for 440799") #+ aes(fill=region)
cat("\n\n")
p + ggtitle("Total trade value for 440799 by region and year
   Vertical red line represents median unit prices (by number of flows for each region), 
   Vertical green line represents aggregate unit prices for each region") + 
    geom_vline(data=swd99region, aes(xintercept = medianprice), colour="red") +
    geom_vline(data=swd99regionunitprice, 
               aes(xintercept = price), colour="green") +
    facet_grid(region + flow ~ year)

# density plot
# ggplot(swd99, aes(price, fill = region)) +
#   geom_density(alpha = 0.2) + xlim(0, 2500) + facet_wrap(~region)

####################################################################### #
# Different way of showing total trade value, using ggplot weight aesthetic
###################################################################### #
p <- ggplot(swd99, aes(x=price, weight=tradevalue)) + 
    geom_histogram(binwidth=100) + xlim(0,2500) +
     ylab("Total trade value in $")
cat("\n\n")
p + facet_grid(flow~.) +
    ggtitle("All world flows for 440799, using price as a weight aesthetic") 

cat("\n\n")
p + facet_grid(region + flow ~ year) +
    ggtitle("Total trade value for 440799 by region and year
            vertical red line represents median price") + 
    geom_vline(data=swd99region, aes(xintercept = medianprice), color="red")
```


### Distribution of unit prices along the tradevalue
Vertical lines, represent large quantities which are at the same prices.
```{r}
swd99$logprice <- round(log(swd99$price))
# Remove infinite prices
cat("\n\n")
p <- ggplot(filter(swd99, price!=Inf),
            aes(x = tradevalue, y = price)) +
    geom_point(alpha = 1/5) +
     coord_cartesian(ylim =c(0,2000)) +
    ggtitle("Variation of prices along the tradevalue")
p + scale_x_log10(limits=c(1000, NA)) + 
    xlab("Trade value $ (log scale)")
p + scale_x_log10(limits=c(1000, NA)) + 
    xlab("Trade value $ (log scale)") + 
    facet_wrap(~flow) + aes(color=as.factor(flag))

summary(swd99[c("tradevalue","price")])

# Zoom on those price levels between first and third quartile
firstquartile <- round(quantile(swd99$price ,probs=0.25, na.rm=TRUE))
thirdquartile <- round(quantile(swd99$price ,probs=0.75, na.rm=TRUE))

p + coord_cartesian(ylim=c(firstquartile,thirdquartile)) +
#     + xlim(c(0,1e6)) +
    scale_x_log10(limits=c(100, NA)) + 
    xlab("Trade value $ (log scale)") + 
    ggtitle(paste("Variation of prices along the trade value
            for price between the first and third quartile",
            firstquartile, "$ and", thirdquartile, "$")) +
    facet_grid(flow + region ~ year) #+ aes(color=year)
# Add a line for median prices on this graph

# Median price and total volume by region
swd99 %>% group_by(flow, region) %>% 
    summarise(tradevalue = sum(tradevalue, na.rm=TRUE),
              price = median(price, na.rm=TRUE)) %>%
    ggplot() + aes(x = tradevalue, y = price, label=region) +
    geom_text() + facet_wrap(~flow) +
    ggtitle("Median prices and total trade value by regions")

# Median price and volume by subregion
swd99 %>% group_by(flow, subregion) %>% 
    summarise(tradevalue = median(tradevalue, na.rm=TRUE),
              price = median(price, na.rm=TRUE)) %>%
    ggplot() + aes(x = tradevalue, y = price, label=subregion) +
    geom_text() + facet_wrap(~flow)
```

### Prices and flags of comtrade estimates
```{r}
swd99 %>% group_by(year, region, flow, flag) %>%
    summarise(medianprice = round(median(price, na.rm=TRUE))) %>%
    dcast(year + region + flow ~ flag, value.var="medianprice" )

```

### Prices of Re-Import and Re-export
```{r}
swd99re <- dtf %>% 
    renamecolumns %>%
    filter(flow %in% c("Re-Import", "Re-Export")) %>%
    merge(select(reportercomtrade, reportercode, region, subregion)) %>%
    mutate(conversion = weight / quantity,
           price = tradevalue / quantity,
           # To avoid "integer overflow - use sum(as.numeric(.))" error 
           # on sum of all values
           tradevalue = as.numeric(tradevalue))
# Remove infinite prices
p <- ggplot(filter(swd99re, price!=Inf),
            aes(x = tradevalue, y = price)) +
    geom_point(alpha = 1/10) +
     coord_cartesian(ylim =c(0,2000)) +
    ggtitle("Variation of prices along the tradevalue")
p + scale_x_log10(limits=c(1000, NA)) + xlab("Trade value $ (log scale)")
p + scale_x_log10(limits=c(1000, NA)) + xlab("Trade value $ (log scale)") + 
    facet_wrap(~flow) + aes(color=year)
```



### TradeValue as a function of the quantity
The slope of those graphs represents unit prices.
```{r}
p <- ggplot(mutate(swd99, splitquantity=round(log(quantity)))) +
    aes(x=quantity, y=tradevalue) + geom_point(alpha = 1/10)
cat("\n\n")
p + scale_x_log10() + scale_y_log10() + 
    ggtitle("Trade value and quantity (log scale)")
cat("\n\n")
p + xlim(c(0,2e3)) + ylim(c(0,1e6)) +
    ggtitle("Trade values below 1e6$, 
            lines represent the median unit price by year and region") + 
    facet_grid(flow+region~year) +
# "             blue line corresponds to a price of 400$/m3,
#             red line to a price of 800$/m3"
#     geom_abline(intercept = 20, slope= 400 , colour = "blue", size = 2) + 
#     geom_abline(intercept = 20, slope= 800 , colour = "red", size = 2)  +
#     geom_smooth(aes(group=splitquantity), method="lm") + 
    geom_abline(data=(swd99region), aes(intercept=0, slope=medianprice))
```

### Linear Model of tradevalue ~ quantity
```{r}
# Very poor fit and irrelevant slope for the whole dataset
mtvq <- lm(tradevalue ~ quantity, swd99)
summary(mtvq)
# Model with log
# mtvqlog <- lm(log(tradevalue)~log(quantity), swd99, na.action=na.exclude)
# For quantities above 500 the slope is a realistic price
mtvq <- lm(tradevalue ~ quantity, filter(swd99, quantity>500& quantity<1000))
summary(mtvq)
```


### TradeValue as a function of the weight

```{r tradevaluefunctionofweight}
p <- ggplot(swd99) + aes(x=weight, y=tradevalue) + geom_point(alpha = 1/10)
#p + ggtitle("Trade value and weight")
# p + xlim(c(0,1e6)) + ylim(c(0,1e6)) + ggtitle("Values below 1e6") + 
#     facet_grid(flow ~ year) + stat_smooth(method="lm")
p + scale_x_log10() + scale_y_log10() + 
    ggtitle("Trade value and weight (log scale)")
```

### Quantity as a function of the weight (conversion factor)
See also the other file on conversion factors.

The slope of those graph represents conversion factors.
```{r}
p <- ggplot(swd99) + aes(x=quantity, y=weight) + geom_point(alpha = 1/10)
# p + ggtitle("Trade value and weight")
cat("\n\n")
p + xlim(c(0,1e3)) + ylim(c(0,1e6)) + ggtitle("Values below 1e6") + 
    facet_wrap(~flow)
cat("\n\n")
p + scale_x_log10() + scale_y_log10() + aes(color=region) +
    ggtitle("Weight and quantity (log scale)") +
    facet_wrap(~region)
```


## Replace quantity in `r nrow(filter(swd99, is.na(quantity)))` flows using unit prices

### One unit prices for each region in each year
Use regional median prices to calculate quantities.
Increase in global trade import and export due to the estimate for all flows
```{r}
# Use regional median prices
swd99regioncorrected <- merge(swd99, swd99region, all.x=TRUE) %>%
    mutate(quantity2 = tradevalue / medianprice,
           havequantity = !is.na(quantity)) 
swd99regioncorrected %>%
    group_by(havequantity,year, flow) %>%
    summarise(rows = n(),
              volume = sum(quantity, na.rm=TRUE) ,
              estimatedvolume = sum(quantity2, na.rm=TRUE),
              increase = estimatedvolume - volume,
              percent = round(increase / volume*100)) %>%
    kable

# Check if both tables contain the same total world trade flows
stopifnot(sum(swd99$quantity,na.rm=TRUE) == 
              sum(swd99regioncorrected$quantity, na.rm=TRUE))


print("Percent change in world trade flows ")
wtf <- sum(swd99$quantity,na.rm=TRUE)
swd99regioncorrected %>% 
    filter(!havequantity) %>% 
    group_by(havequantity, year, flow) %>%
    summarise(percentchangeinworldtrade = 
                  round((sum(quantity2,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / 
                            wtf*100,3))
```


### One unit prices for each subregion in each year
```{r}
# Use  sub-regional median prices
swd99subregioncorrected <- merge(swd99, swd99subregion) %>%
    mutate(quantity3 = tradevalue / medianprice,
           havequantity = !is.na(quantity)) %>%
    group_by(havequantity, year, flow) %>%
    summarise(rows = n(),
              volume = sum(quantity, na.rm=TRUE) ,
              estimatedvolume = sum(quantity3, na.rm=TRUE),
              increase = estimatedvolume - volume,
              percent = round(increase / volume*100)) 
swd99subregioncorrected %>% kable

print("Total world trade increase")
# Check if both tables contain the same total world trade flows
stopifnot(sum(swd99$quantity,na.rm=TRUE) == sum(swd99subregioncorrected$volume))

print("Percent change in world trade flows ")
swd99subregioncorrected %>% filter(!havequantity) %>% 
    group_by(year, flow) %>%
    summarise(percentchangeinworldtrade = 
                  round(sum(increase)/sum(swd99$quantity,na.rm=TRUE)*100,3))
```


## Replace quantities when unit prices are too high or too low
### Unit prices are too high
```{r}
swd99regioncorrected %>% filter(price > upperprice ) %>%
    group_by(year, flow) %>% 
    summarise(percentchangeinworldtrade = 
                  round((sum(quantity2,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / 
                            wtf*100,3)) %>%
    kable
```

Double the threshold for upperprice
```{r}
swd99regioncorrected %>% filter(price > 2*upperprice ) %>%
    group_by(year, flow) %>% 
    summarise(percentchangeinworldtrade = 
                  round((sum(quantity2,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / 
                            wtf*100,3)) %>%
    kable
```

### Unit prices are too low
```{r}
swd99regioncorrected %>% filter(price < lowerprice ) %>%
    group_by(year, flow) %>% 
    summarise(percentchangeinworldtrade = 
                  round((sum(quantity2,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / 
                            wtf*100,3)) %>%
    kable
```

Half the threshold for lower price
```{r}
swd99regioncorrected %>% filter(price < lowerprice/2 ) %>%
    group_by(year, flow) %>% 
    summarise(percentchangeinworldtrade = 
                  round((sum(quantity2,na.rm=TRUE) - 
                             sum(quantity,na.rm=TRUE)) / 
                            wtf*100,3)) %>%
    kable
```


