---
title: "Data completeness by country for product ` as.character(productcodeinreport)` "
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
pdf_document: >
  \usepackage[utf8]{inputenc}
---


This is the development version of the completeness report.
The version distributed with the package is under
inst/templates

## Warning
This report is generated automatically based on data from the 
[UN Comtrade API](http://comtrade.un.org/data/Doc/api/ex).
Tables and graphs are in development.
They are intended for internal use within the TradeFLowsDB project.
Country numbers in tables below might be slightly
overestimated, use them as a rough indication only.
```{r loaddata}
library(dplyr)
library(tradeflows)
library(reshape2)
library(ggplot2)
library(knitr)
```


```{r prepare_completeness_table, eval=TRUE}
tfdata <- readdbproduct(440799, "raw_flow_yearly")
productcodeinreport <- 440799

# This report is intended for one product only
tfdata2 <- tfdata %>% 
    addregion %>%
    filter(productcode == productcodeinreport) %>%
    filterworldeu28

completeness <- tfdata2 %>%
    select(reporter, classification, year, quantity, weight, tradevalue) %>%
    melt(id=c("reporter", "classification", "year"))  %>%
    mutate(value = !is.na(value)) %>%
    unique %>% 
    arrange(classification, reporter, year)
productcodeinreport <- as.character(productcodeinreport)
```


```{r prepareflagtable}
flagcountry <- tfdata2 %>%
    group_by(reporter, year, flow, flag) %>%
    filter(flow %in% c("Import", "Export")) %>%
    summarise(tradevalue = sum(tradevalue),
              quantity = sum(quantity, na.rm = TRUE),
              weight = sum(weight, na.rm = TRUE)) %>%
    data.frame %>% # Needed otherwise it will return factors with only one level Factor w/ 1 level "0":
    mutate(flag = as.factor(flag))

flagregion <- tfdata2 %>%
    group_by(regionreporter, year, flow, flag) %>%
    filter(flow %in% c("Import", "Export")) %>% # Remove re-import and re-export
    summarise(tradevalue = sum(tradevalue),
              quantity = sum(quantity, na.rm = TRUE),
              weight = sum(weight, na.rm = TRUE)) %>% 
    data.frame %>% # Needed otherwise it will return factors with only one level Factor w/ 1 level "0":
    mutate(flag = as.factor(flag))

str(flagregion$flag)
```


## Product definitions
Product `r productcodeinreport` has the following definitions:
```{r productdefinition, results='asis', eval=TRUE}
productname <- tfdata2 %>% 
    select(classification, productdescription) %>%
    arrange(classification) %>%
    unique

for(code in productname$classification){
    cat("\n\n__",code,":__ ", sep="")
    cat(productname$product[productname$classification == code])
}
```


## Summary
Number of reporting countries by year and by classification type.
```{r reporting_by_classification}
completeness %>% 
    select(reporter, year, classification) %>%
    unique() %>% # Keep only unique combinations of reporter, year, classification
    group_by(year, classification) %>% 
    dcast(year ~ classification, value.var="reporter", length) %>%
    kable()
```

List of countries using each classification type in `r max(completeness$year)`.
```{r reportername_by_classification, results='asis', eval=TRUE}
reporterbyclassif <- completeness %>% 
    filter(year == max(year)) %>%
    select(year, classification, reporter) %>%
    unique()
#     group_by(classification) %>%
#     do(head(.,2)) # works! how to use this do

for (code in unique(reporterbyclassif$classification)){
    cat("\n\n__",code,":__ ", sep="")
    cat(reporterbyclassif$reporter[reporterbyclassif$classification == code], 
        sep=", ")
}

```

## Completeness by country and by classification type 
```{r plot, results='asis', eval=TRUE}
# Prepare table to show partner countries
# remove world
tfdata3 <- tfdata2 %>% 
    filterworldeu28 %>%
    select(year, classification, flow, reporter, partner) %>%
    arrange(year, classification, flow)
for (r in unique(completeness$reporter)){
    cat("\n\n###", r, "\n")
    
    # plot completeness
    p <- ggplot(data=filter(completeness, reporter == r), 
                aes(x = year, y = variable, 
                    xend = year + 1, yend = variable, 
                    color = value, linetype = value)) +
        scale_colour_manual(values = c("TRUE" = "blue", "FALSE" = "red")) +
        scale_linetype_manual(values = c("TRUE" = 2, "FALSE" = 1)) +
        geom_segment(size=3) + 
        facet_grid(classification~.) +
        ggtitle(paste("Data completeness of", productcodeinreport,
                      "in", r))
    plot(p)

    
    ###################################################### #
    # Number of trade partners per classification and year #
    ###################################################### #
    # Reported by the country
    npartnersbyr <- tfdata3 %>% 
        filter(reporter == r) %>%
        dcast(year + classification ~ flow, value.var="partner", length) 
    try(print(kable(npartnersbyr,
                    caption = paste("Number of trade partners reported by", r))))
    
    # Reported by partner countries
    try(npartnersbyothers <- tfdata3 %>% 
            filter(partner == r) %>%
            dcast(year + classification ~ flow, value.var="reporter", length))
    try(print(kable(npartnersbyothers,
                    caption = paste("Number of countries who report trading with",
                                    r))))
    }
``` 



```{r dev_completenessforFrance}


r = "France"
# plot completeness
p <- ggplot(data=filter(completeness, reporter == r), 
            aes(x = year, y = variable, 
                xend = year + 1, yend = variable, 
                linetype = value)) +
    scale_colour_manual(values = c("TRUE" = "blue", "FALSE" = "red")) +
    scale_linetype_manual(values = c("TRUE" = 2, "FALSE" = 1)) +
    geom_segment(aes(color = value), size=3) + 
    geom_text(aes(y = "classification", label=classification)) +
#     coord_cartesian(ylim = c(-0.3, 3.5)) +
    ggtitle(paste("Data completeness of", productcodeinreport,
                  "in", r))
plot(p)

createreportfromdb
# plot flag ##########
pflagfr <- ggplot(data = filter(flagcountry, reporter == r),
                aes(x= year, y = tradevalue, fill = flag)) + 
    geom_bar(stat = "identity") + facet_grid(flow ~ ., scales = "free_y") 
plot(pflagfr)
pflagfr + aes(y = weight)
pflagfr + aes(y = quantity)

message("Change flag labels to a short description for the plot")

pflag <- ggplot(data = flagregion,
                aes(x = year, y = tradevalue, fill = flag)) + 
    geom_bar(stat = "identity") +
    facet_grid(flow ~ regionreporter, scales = "free_y")
plot(pflag)
pflag + aes(y = weight)
pflag + aes(y = quantity)


# what are those extreme quantities?
bigquantity <-filter(tfdata2, quantity>1e6) %>%
    mutate(flag = paste("Flag", flag))
ggplot(data = bigquantity,
       aes(x = reporter,y = quantity, label = partner, 
           color = as.factor(year))) +
    geom_text() + scale_y_log10() + 
    facet_grid(flag ~ flow) + theme_bw()

ggplot(data = bigquantity,
       aes(x = year, y = quantity, fill = as.factor(flag))) + 
    geom_bar(stat = "identity") +
    facet_grid(flow ~ regionreporter, scales = "free_y") +
    ggtitle("It seems all large quantities are flag 0 or 4")

ggplot(data = filter(tfdata2, quantity<1e7),
       aes(x = year, y = quantity, fill = as.factor(flag))) + 
    geom_bar(stat = "identity") +
    facet_grid(flow ~ regionreporter, scales = "free_y") +
    ggtitle("quantity < 1e7")

```
