---
title: "Country name encoding in utf-8 only"
author: "Paul Rougieux"
date: "26/02/2015"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=TRUE, warning=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 



## Simo's message from August 10
Simo:

> the first validation did not produce same number of rows:

> 37415 rows in the raw table became 33280 rows in the validated table.

> The attachment is that raw table.

## Investigating fuel wood data
I renamed "raw_flow_yearly" to "raw_flow_fuelwood" and 
Loaded the raw table into Mysql with the command:
```
cat raw_flow_fuelwood.sql | mysql -u paul -p tradeflows
```

```{r}
fw <- readdbtbl("raw_flow_fuelwood")
fwprod <- fw %>% group_by(productcode) %>%
    summarise(nrow = n()) %>% 
    arrange(desc(nrow)) %>% data.frame()
fwprod %>% kable
# Are there duplicates?
fwdall <- fw %>% collect()
findduplicatedflows(fwdall)
```

## Cleaning fuel wood data
```{r}
# Add column to the column definitions to avoid complains
column_names$raw_flow_fuelwood <- column_names$raw_flow_yearly
# Clean one of the product in this table
cleandbproduct(440130, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
cleandbproduct(440110, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
cleandbproduct(440122, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
cleandbproduct( 440121, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
cleandbproduct( 440320, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
cleandbproduct(440290, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
cleandbproduct(440200, tableread =  "raw_flow_fuelwood", tablewrite = "validated_flow_yearly")
```

## Compare number of rows in the raw and validated datasets
```{r}
fwraw <- readdbtbl("raw_flow_fuelwood") %>%
    group_by(productcode) %>%
    summarise(raw_lines = n()) %>% 
    arrange(desc(raw_lines)) %>% data.frame()

fwvalidated <- readdbtbl("validated_flow_yearly") %>%
    filter(productcode %in% fwraw$productcode) %>%
    group_by(productcode) %>%
    summarise(validated_lines = n()) %>% 
    arrange(desc(validated_lines)) %>% data.frame() %>%
    merge(fwraw)
fwvalidated %>% kable

```

| productcode|  nrow|
|-----------:|-----:|
|      440130| 13763|
|      440110|  7317|
|      440122|  7232|
|      440121|  4968|
|      440320|  1507|
|      440290|  1379|
|      440200|  1249|


no reim port and re export
```{r}
readdbtbl("raw_flow_fuelwood") %>%
    filter(flow %in% c("Import","Export")) %>%
    group_by(productcode) %>%
    summarise(raw_lines = n()) %>% 
    arrange(desc(raw_lines)) %>% data.frame()


``` 


