---
title: "440349 China Malaysia"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tradeflows)
library(knitr)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(dplyr)
library(ggplot2)
library(tidyr)
``` 


# Mail discussion December 2015

Here is the issue mentioned by Jean from ITTO in his December 2015 email :

> 3) Querying the data: When I query 2005-China-volume
> import-from Malaysia-code 440349, I get roughly 370 000 m3.
> When I query the mirror stat I get something like 195 000 m3. 
> My understanding of this project was somehow that
> the system would give us
> what it thinks it is the best estimate for this trade flow.

[Suggestion to suppress RMySQL's lengthy warnings concerning data type changes](https://github.com/rstats-db/RMySQL/issues/37).
    

# 2005 China Malaysia 440349 trade discrepancy issue
```{r}
logs49r <- suppressWarnings(readdbproduct(440349, "raw_flow_yearly")) %>%
    filterworldeu28()
logs49v <- suppressWarnings(readdbproduct(440349, "validated_flow_yearly"))
min(logs49r$year)

chimal <- c("China", "Malaysia")
bilflows <- function(dtf, countries = c("China", "Malaysia"), minyear=0, maxyear=9999){
    dtf %>% 
        filter(year >= minyear & year <= maxyear) %>%
        filter(reporter %in% countries & partner %in% countries) %>%
        select(productcode, flow, year, reporter, partner, tradevalue, quantity, flag)
}
logs49r %>% bilflows(minyear=2005, maxyear=2005) 
logs49v %>% bilflows(minyear=2005, maxyear=2005) 

logs49rcleaned <- logs49r %>% clean()
logs49rcleaned %>% bilflows(minyear=2005, maxyear=2005)
```


# Try to correct the issue on 2010 data
Update 14 December: the issue has also been fixed on 2005 data now.
Before the change, the 2 tables below illustrate discrepancies remaining in the cleaned data.
```
# Raw data
productcode   flow year reporter  partner tradevalue quantity flag
1      440349 Import 2010    China Malaysia   84356413   391076    0
2      440349 Export 2010 Malaysia    China   28229869   185550    4

# validated data EFI
  productcode   flow year reporter  partner tradevalue quantity flag
1      440349 Import 2010    China Malaysia   84356413   185550 4000
2      440349 Export 2010 Malaysia    China   28229869    29223  304

# Cleaned data on my system (slight difference due to change in price rounding)
  productcode   flow year reporter  partner tradevalue  quantity flag
1      440349 Import 2010    China Malaysia   84356413 185550.00 4000
2      440349 Export 2010 Malaysia    China   28229869  29230.76  304

# The difference is due to the fact that replacebypartnervalue() is 
# called before shaveprice()
# I changed the order in which those 2 functions are called.
```


```{r}
cleaned <- logs49r %>% clean()

logs49v %>% bilflows(minyear=2010, maxyear=2010)
# Raw data
logs49r %>% bilflows(minyear=2010, maxyear=2010)
cleaned %>% bilflows(minyear=2010, maxyear=2010)

logs49r %>% 
    filter(year == 2010) %>%
    filter(reporter %in% chimal & partner %in% chimal) %>%
    addpartnerflow() %>%
    addconversionfactorandprice() %>%
    select(productcode, flow, year, reporter, partner, reportercode, partnercode, 
           tradevalue, quantity, flag, quantitypartner)
```


# Illustrate price variation
```{r}
chimaldata <- logs49r %>%
    filter(reporter %in% chimal & partner %in% chimal) %>%
    addconversionfactorandprice() 

ggplot(chimaldata, aes(x=year, y=price, color=reporter)) + geom_point() +
    facet_wrap(~flow) + ggtitle("prices of 440349 in China and Malaysia") + ylim(0,NA)

message("Variables that identify a unique flow")
flowidvariables <- c("year", "reporter","partner", "flow","productcode")
# chimaldata2 <- gather(chimaldata, key, value, -year, -reporter, -partner, -flow, -productcode),
#        aes(x = year, y = value, 
# ggplot()

chimaldata %>% select(year, reporter, partner, flow, tradevalue, quantity, price) %>% kable()
```


# 440320 issues with NA quantity
```{r}
devtools::load_all() # Load unexported function replacebypartnerquantity()
log20r <- readdbproduct(440320,"raw_flow_yearly") 
log20v <- readdbproduct(440320,"validated_flow_yearly")
log20vna <- log20v %>% filter(is.na(quantity))
chnrus <- c("China","Russian Federation")
log20r %>% bilflows(chnrus, minyear = 2014, maxyear = 2014)
log20v %>% bilflows(chnrus, minyear = 2014, maxyear = 2014)
log20vna %>% bilflows(chnrus)
fraswe <- c("France", "Sweden")
log20r %>% bilflows(fraswe, minyear = 2012, maxyear = 2012)
log20v %>% bilflows(fraswe, minyear = 2012, maxyear = 2012)
# Raw data has quantity for both flows, validated data has a missing quantity
# Add a breakpoint in the replacebypartnerquantity() function

# This data frame contains data for France and Sweden, and I want to mirror it
fs <- log20r %>% 
    filter(reporter %in% fraswe & partner %in% fraswe & year ==2012) %>%
    mutate(tradevaluepartner = NA, quantitypartner = NA)
choicefs <- fs %>% select(flow, reportercode, partnercode) %>% 
    mutate(favorpartner = c(FALSE, FALSE, TRUE, FALSE)) 
# A NA value should not appear under quantitypartner
# There is an issue with the merge
# The merge should be done at some point
fs %>%  replacebypartnerquantity(choicefs)
fs %>% findduplicatedflows()
```

# Flag 5000 missing?
```{r}
dtf <- readdbtbl("validated_flow_yearly")
dtf %>% group_by(flag) %>%
    summarise(number = n(),
              quantity = sum(quantity)) %>% kable()

prodcode <- dtf %>% select(productcode) %>% distinct() %>% collect()
prodcode$productcode

dtfr <- readdbtbl("raw_flow_yearly") %>%
    select(productcode) %>% distinct() %>% collect()
dtfr <- dtfr %>% mutate(digit = round(productcode/1e2)) 
unique(dtfr$digit)

# Clean a dataset to see
swd99r <- readdbproduct(440799, "raw_flow_yearly")
swd99v <- clean(swd99r)
unique(swd99v$flag)
```

# Flag 9000 shouldn't appear
```
swd99r %>% bilflows(bulden, minyear = 2009, maxyear = 2009)
 productcode   flow year reporter  partner tradevalue quantity flag
1      440799 Import 2009  Denmark Bulgaria       1511        5    2

swd99v %>% bilflows(bulden, minyear = 2009, maxyear = 2009)
  productcode   flow year reporter  partner tradevalue quantity flag
1      440799 Export 2009 Bulgaria  Denmark       1511       NA 9302
2      440799 Import 2009  Denmark Bulgaria       1511       NA 4302
```

```{r}
# BUlgaria Denmark 2009 440799
bulden <- c("Bulgaria", "Denmark")
swd99r %>% bilflows(bulden, minyear = 2009, maxyear = 2009)
swd99v %>% bilflows(bulden, minyear = 2009, maxyear = 2009)
```
