---
title: "Geneva Meeting February 2016"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tradeflows)
library(knitr)
opts_chunk$set(message=FALSE, warning = FALSE)
opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(dplyr)
library(ggplot2)
library(tidyr)
``` 


# 2005 China Malaysia 440349 trade discrepancy issue
```{r}
chmy <- readdbtbl("validated_flow_yearly") %>%
    filter(reporter == "China" & partner == "Myanmar") %>%
    collect()
mych <-  readdbtbl("validated_flow_yearly") %>%
    filter(reporter == "Myanmar" & partner == "China") %>%
    collect()

chmyr <- readdbtbl("raw_flow_yearly") %>%
    filter(reporter == "China" & partner == "Myanmar") %>%
    collect()
mychr <-  readdbtbl("raw_flow_yearly") %>%
    filter(reporter == "Myanmar" & partner == "China") %>%
    collect()
```


# 440724 
```{r }
# Find which product has a high volume issue in the validated data
cam <- readdbtbl("validated_flow_yearly") %>%
     filter((reporter == "Cameroon" | partner == "Cameroon") & 
                unit == "Volume in cubic meters") %>%
     arrange(desc(quantity)) %>%
     head(200)

#' Display a sorted table for the purpose of this analysis
#' @param dtf a data frame of trade flows
littletable <- function(dtf,n=2){
    dtf %>% 
        select(reporter, partner, year, flow, tradevalue, quantity, weight, price) %>%
        arrange(desc(quantity)) %>%
        head(n) %>%
        kable()
}
```


## Raw data shows Kuwait, Botswana and South Africa import extreme quantities
```{r}
sawn24r <- readdbproduct(440724, "raw_flow_yearly") %>% 
    filterworldeu28() %>%
    addconversionfactorandprice() %>%
    addregion()
sawn24r %>% littletable()
sawn24r %>% filter(regionreporter == "Africa") %>% littletable()
```


## Validated data shows Cameroon Netherlands with extreme volume and price issue
Kuwait, Botswana and South Africa trade flows were corrected in the validated data,
but not the export flows between Cameroon and Netherlands remained erroneous
```{r}
sawn24v <- readdbproduct(440724, "validated_flow_yearly") %>%
    addconversionfactorandprice() %>%
    addregion()
sawn24v %>% littletable()
```

Archive table to compare once the algorithm has been modified

|reporter |partner     | year|flow   | tradevalue| quantity|   weight|   price|
|:--------|:-----------|----:|:------|----------:|--------:|--------:|-------:|
|Cameroon |Netherlands | 2007|Export |   30350149|  9547163| 52119188| 3.17897|
|Cameroon |Netherlands | 2008|Export |   37635852|  5136247| 57231874| 7.32750|


## Price extraction issues
Analyse 2007 exports from Africa. Only 18 from 120 entries have a price. All other values are  not available (NA). NA values should not be taken into account in the calculation
```{r eval=TRUE}
sawn24r2007afr <- sawn24r %>% filter(regionreporter == "Africa" & year == 2007)
sawn24r2007afr %>% 
    mutate(value = ifelse(is.na(price),"NA","value"),
           value = ifelse(is.infinite(price),"Inf",value)) %>%
    group_by(value) %>% summarise(nrow = n()) %>% kable()

# Change in median price when removing NA values, same as option na.rm
median(sawn24r2007afr$price)
median(sawn24r2007afr$price, na.rm = TRUE)
median(sawn24r2007afr$price[!is.na(sawn24r2007afr$price)])

# Change in median price when removing NA and Inf values
median(sawn24r2007afr$price[!is.na(sawn24r2007afr$price) & !is.infinite(sawn24r2007afr$price)])

price <-  sawn24r %>%
    extractprices()

ggplot(price, aes(x = year, y = medianprice, ymin = lowerprice, ymax = upperprice)) +
    geom_point() + geom_errorbar() +
    facet_grid(unit + flow ~ regionreporter) +
    ggtitle("Regional median prices and price bounds")

ggplot(price, aes(x = year, y = medianprice, color = regionreporter)) +
    geom_line() + facet_grid(flow ~ .) +
    ggtitle("Global median prices")

priceglobal <- sawn24r %>%
    extractprices(grouping = c("flow", "year", "unit"))

ggplot(priceglobal, aes(x = year, y = medianprice, ymin = lowerprice, ymax = upperprice)) +
    geom_point() + geom_errorbar() +
    facet_grid(unit + flow ~ . ) + 
    ggtitle("Global median prices and price bounds")

ggplot(priceglobal, aes(x = year, y = medianprice, color = flow)) +
    geom_line() + 
    ggtitle("Global median prices")

# Try smooth
plot(priceglobal$medianprice[priceglobal$flow == "Import"])
plot(smooth(priceglobal$medianprice[priceglobal$flow == "Import"]))
plot(priceglobal$medianprice[priceglobal$flow == "Export"])
plot(smooth(priceglobal$medianprice[priceglobal$flow == "Export"]))
priceglobal <- priceglobal %>% 
    group_by(flow) %>%
    mutate(mediansmooth = as.numeric(smooth(medianprice)))
ggplot(priceglobal, aes(x = year, y = mediansmooth, color = flow)) +
    geom_line() + 
    ggtitle("Global median prices")
```

    
## Correct the "shave price" algorithm
Place a breakpoint in the price calculation to see the price data
```{r}
sawn24c <- sawn24r %>% clean()
# Number of rows and price distribution
# This can be used to export prices and conversion factors to an excel file
# cleandb2excel(440724)
```


## Conversion factor issues
Infinite conversion factor should not be taken into account in the 
calculation of the mean conversion factor.
```{r}
conv <- sawn24r %>% 
    extractconversionfactors()
sum(conv$medianconversion)

ggplot(conv, aes(x = year, y = medianconversion)) +
    geom_point() + 
    facet_grid(unit + flow ~ regionreporter )
```


# Comtrade flags in the raw data 
```{r}
nrowflag <- readdbtbl("raw_flow_yearly") %>%
    select(flag, unit) %>%
    group_by(flag, unit) %>%
    summarise(n = n()) %>% collect()
```


## Include comtrade estimates in the price calculation
Functions that extract prices and conversion factors have flag
that can disable comtrade estimates if required. 
This is not activated by default.
