---
title: "Data completeness by country for product `r as.character(productcodeinreport)` "
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
pdf_document: >
  \usepackage[utf8]{inputenc}
---


## Warning
This report is generated automatically based on data from the 
[UN Comtrade API](http://comtrade.un.org/data/Doc/api/ex).
Tables and graphs are in development.
They are intended for internal use within the TradeFLowsDB project.
Country numbers in tables below might be slightly
overestimated, use them as a rough indication only.


```{r generatereportdatafordevelopment, eval=FALSE}
# Generate report data for development purposes only
tfdata2 <- readdbproduct(440799, "raw_flow_yearly")
# optional filter by year
#tfdata2 <- tfdata2 %>% filter(year %in% c(2010,2011))

productcodeinreport <- 440799
# Then generate the completeness data frame below
# These data frames should be enough to run all chunks below
# Give a value to r to generate only one of the completeness plot 
# in the loop, in the plot chunk below
```


```{r prepare_completeness_table, eval=TRUE}
# This report is intended for one product only
tfdata2 <- tfdata %>% 
    filter(productcode == productcodeinreport) %>%
    filterworldeu28

completeness <- tfdata2 %>%
    select(reporter, classification, year, quantity, weight, tradevalue) %>%
    melt(id=c("reporter", "classification", "year"))  %>%
    mutate(value = !is.na(value)) %>%
    unique %>% 
    arrange(classification, reporter, year)
productcodeinreport <- as.character(productcodeinreport)
```


## Product definitions
Product `r productcodeinreport` has the following definitions:
```{r productdefinition, results='asis', eval=TRUE}
productname <- tfdata2 %>% 
    select(classification, productdescription) %>%
    arrange(classification) %>%
    unique

for(code in productname$classification){
    cat("\n\n__",code,":__ ", sep="")
    cat(productname$product[productname$classification == code])
}
```


## Summary
Number of reporting countries by year and by classification type.
```{r reporting_by_classification}
completeness %>% 
    select(reporter, year, classification) %>%
    unique() %>% # Keep only unique combinations of reporter, year, classification
    group_by(year, classification) %>% 
    dcast(year ~ classification, value.var="reporter", length) %>%
    kable()
```


List of countries using each classification type in `r max(completeness$year)`.
```{r reportername_by_classification, results='asis', eval=TRUE}
reporterbyclassif <- completeness %>% 
    filter(year == max(year)) %>%
    select(year, classification, reporter) %>%
    unique()
#     group_by(classification) %>%
#     do(head(.,2)) # works! how to use this do

for (code in unique(reporterbyclassif$classification)){
    cat("\n\n__",code,":__ ", sep="")
    cat(reporterbyclassif$reporter[reporterbyclassif$classification == code], 
        sep=", ")
}
```


## Completeness by country and by classification type 
```{r plot, results='asis', eval=TRUE}
# Prepare table to show partner countries
# remove world
tfdata3 <- tfdata2 %>% 
    filterworldeu28 %>%
    select(year, classification, flow, reporter, partner) %>%
    arrange(year, classification, flow)
years <- unique(tfdata3$year) # for the scales
for (r in unique(completeness$reporter)){
    cat("\n\n###", r, "\n")
    
    # plot completenes
    p <- ggplot(data=filter(completeness, reporter == r), 
                aes(x = year, y = variable, 
                    xend = year + 1, yend = variable, 
                    linetype = value)) +
        scale_colour_manual(values = c("TRUE" = "blue", "FALSE" = "red")) +
        scale_linetype_manual(values = c("TRUE" = 2, "FALSE" = 1)) +
        scale_x_continuous(breaks = years) +
        geom_segment(aes(color = value), size=3) + 
        geom_text(aes(y = "classification", label=classification)) +
        ggtitle(paste("Data completeness of", productcodeinreport,
                      "in", r))
    plot(p)

    ###################################################### #
    # Number of trade partners per classification and year #
    ###################################################### #
    # These try statements have been introduced because there is sometimes missing
    # data. It would be nice to deal with these errors and print something when they occur
    # Reported by the country
    npartnersbyr <- tfdata3 %>% 
        filter(reporter == r) %>%
        dcast(year + classification ~ flow, value.var="partner", length) 
    try(print(kable(npartnersbyr,
                    caption = paste("Number of trade partners reported by", r))))
    
    # Reported by partner countries
    try(npartnersbyothers <- tfdata3 %>% 
            filter(partner == r) %>%
            dcast(year + classification ~ flow, value.var="reporter", length))
    try(print(kable(npartnersbyothers,
                    caption = paste("Number of countries who report trading with",
                                    r))))
}
``` 
