---
title: '`r as.character(productcodeinreport)` discrepancies in `r reporterinreport`'
author: "Paul Rougieux"
date: "03/12/2014"
output:
  pdf_document:
    fig_height: 7
    fig_width: 10
    toc: no
    toc_depth: 3
---

```{r packages, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
# I tried placing the yaml block after the chunks but for some reason it doesn't always generate the pdf and generates a html file instead.
# pdf is generated when the yaml block is at the top, 

library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 


```{r development, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
# Remove this chunk (or change eval=FALSE) when this file is used as a template
# The data frame tfdata will be passed by the report generating function to the template
# Product and reporter information can be extracted back from tfdata 
# in the template this would have to be the country appearing most in the data frame
productcodeinreport <- c(440799)
reporterinreport <- "Cameroon" 
tfdata <- readdbtbl("raw_flow_yearly") %>%
    filter(productcode == productcodeinreport & (reporter == reporterinreport | partner == reporterinreport)) %>%
    collect
```


```{r prepareproductandreporter}
metadata <- extractmetadata(tfdata)
stopifnot(metadata$productcode == productcodeinreport)
stopifnot(metadata$reporter == reporterinreport)
if(metadata$unit == "Volume in cubic meters") metadata$unit <- "cubic meters"
# Calculate discrepancies
tfdata  <- tfdata %>% addpartnerflow %>% calculatediscrepancies 

# Find the last year for which the reporter reported data
lastyearreporter <- tfdata %>% filter(reporter == reporterinreport) %>% filter(year==max(year)) %>%
    select(year) %>% unique
lastyearreporter <- lastyearreporter$year
# The last year in the dataset tends to have weak data, take lastyear - 2 as the report year.
# But if lastyearreporter is still smaller than that take it anyway
yearinreport <- min(as.numeric(lastyearreporter), max(as.numeric(tfdata$year))-2)

# Display number of years available only if other countries reported 
# data for more years
if (lastyearreporter != max(tfdata$year)){
    messageyears = paste("Data for other countries is available until",
                         max(tfdata$year),".")
} else {
    messageyears = ""
}
    
```

`r reporterinreport` reported importing from
`r length(unique(tfdata$partner[tfdata$reporter==reporterinreport & tfdata$flow=="Import"]))` countries and exporting to
`r length(unique(tfdata$partner[tfdata$reporter==reporterinreport & tfdata$flow=="Export"]))` countries 
between `r min(tfdata$year)` and `r lastyearreporter`.
`r messageyears`

```{r numberofreportedflows}
tfdata %>% 
    mutate(isreporter = ifelse(reporter == reporterinreport,
                               paste("flows reported by",
                                     reporterinreport),
                               "reported by its partners")) %>%
    group_by(year,flow, isreporter) %>%
    summarise(nrow = n()) %>%
    dcast(year + flow ~ isreporter,value.var = "nrow") %>%
    kable
```


\newpage

## 9 largest trade quantities in `r yearinreport` 
 9 largest import or export flows reported by `r reporterinreport`. 
 Trade value is reported in US dollars and quantity in `r metadata$unit`.

```{r}
# Select the largest import or export flows reported by the reporter
largestreporter <- tfdata %>% 
    filterworldeu28 %>%
    filter(year == (max(as.integer(year)) -2) & 
               reporter == reporterinreport) %>% 
    arrange(desc(quantity)) %>%
    head(9)

# Display latest trade flows as a table
largestreporter %>% 
    select(year, reporter, partner, flow, flag, tradevalue, quantity,
           dicrepancy_quantity = discrq) %>%
    kable

# Display these trade flows as a plot
largestreporter <- tfdata %>% 
    filterworldeu28 %>% 
    # Keep the same combination of flow country
    filter(paste0(flow,partner) %in% 
               paste0(largestreporter$flow,
                      largestreporter$partner)) %>%
    # reshape table for use in the plot
    select(year, partner, flow, quantityreporter, quantitypartner) %>%
    arrange(-quantityreporter) %>%
    rowwise() %>%
    mutate(qmin = min(quantityreporter, quantitypartner),
           qmax = max(quantityreporter, quantitypartner)) %>%
#     rename(discrepancy = discrq) %>%
    melt(id = c("year","partner","flow", "qmin","qmax"))

ggplot(largestreporter, aes(x = year, y = value, fill = variable)) +
    geom_bar(position="dodge", stat="identity") +
    scale_fill_manual(values = c("chocolate3","black"))+
    geom_errorbar(aes(ymin = qmin, ymax = qmax, 
                      width=0.2, colour="discrepancy")) +
    scale_colour_manual(name="error bars",values="red") +
    ylab(metadata$unit) +
    facet_wrap(flow~partner) +
    theme_bw() + theme(legend.position= "bottom",
                       legend.box="horizontal") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r eval=FALSE}
## Trade flows as a point cloud
tfdata <- tfdata %>% filterworldeu28 %>% addregion
ggplot(tfdata, aes(x = year, y = quantity, color = regionreporter)) +
    geom_point() + facet_wrap(~flow) 
```

