---
title: "Overview of HS codes"
author: "Paul Rougieux"
date: "11 November 2015"
header-includes:
   - \usepackage{longtable}
output: 
  pdf_document: 
    toc: yes
---

```{r message=FALSE}
library(tradeflows)
library(dplyr)
library(tidyr)
library(knitr)
library(xtable)
options(xtable.comment = FALSE) 
```


# An overview of changes in HS codes 
This is Jo's question:

> A big thing in terms of pages is that I would like to replace the ANNEX a1 with an overview of the HS codes and the HS system under which they are applicable.

I think he wants to display changes in HS codes 
between H1 and H2, between H2 and H3 and between H3 and H4.
There are much less changes between H3 and H4 than between the other ones.


```{r results='asis'}
#' Print a table of productcode, description, parentcode, with nice formating 
#' @examples printtable(mutate(cars,bli = speed+1))
printtable <- function(dtf){
    dtf %>%
        xtable(align = "p{50pt}p{50pt}p{300pt}p{55pt}") %>%
        print.xtable(tabular.environment = "longtable", 
                     # booktabs = TRUE,
                     include.rownames=FALSE, type = "latex", floating=FALSE)
}
message("use anti_joint instead")
#' Display product codes that disappear and product codes that appear during a change of 
#' HS classification
#' @param Ha old classification
#' @param Hb new classification
#' @param newerHSname name of the newer HS classification
changes <- function(Ha,Hb,newerHSname){
    # Check that description is non-empty 
    # missing description will be used below after the merge 
    # to check codes appearing and disapearing 
    stopifnot(sum(is.na(Ha$description))==0)
    stopifnot(sum(is.na(Hb$description))==0)
    HSchanges <- full_join(Ha,Hb,by="productcode")
    cat("\n\n## Product codes that disappeared in",newerHSname,"\n\n")
    # cat(HSchanges$productcode[is.na(HSchanges$description.x)], sep=", ")
    HSchanges %>%
        filter(is.na(description.y)) %>%
        select(productcode, description.x, parentcode.x) %>%
        printtable()

    
    cat("\n\n## Product codes that appeared in",newerHSname,"\n\n")
    # cat(HSchanges$productcode[is.na(HSchanges$description.x)], sep=", ")
    HSchanges %>% 
        filter(is.na(description.x)) %>% 
        select(productcode, description.y,parentcode.y) %>%
        printtable()
}
```


\newpage

```{r listofchanges, results='asis'}
attach(classificationcomtrade)
changes(H1,H2,"H2")
cat("\n\n\\newpage")
changes(H2,H3,"H3")
cat("\n\n\\newpage")
changes(H3,H4,"H4")
detach(classificationcomtrade)
anti_join(classificationcomtrade$H3,classificationcomtrade$H4,by="productcode")
```

