---
title: "Interaction with the Tradeflows Database and PHP code"
author: "Paul Rougieux"
date: "26/02/2015"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3
---


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_knit$set(root.dir="../../..") # file paths are relative to the root of the project directory
opts_chunk$set(echo=TRUE, warning=FALSE)
library(tradeflows)
library(dplyr)
library(ggplot2)
library(reshape2)
``` 

```{r loadandprepare}
rawflow0 <- readdbproduct(440799, "raw_flow_yearly")
rawflow1 <- readdbproduct(440799, "raw_flow_year")
nrow(rawflow0)
nrow(rawflow1)
```



## 2015 February 26 email discussion

Paul:

> It seems we do not have the same content in raw_flow_yearly, your
> output says 20996 rows in the dataset while mine says 25208 rows

Simo:

> Maybe the Comtrade server is not behaving consistently.. I tried to "refresh" the data earlier, no change in the row count. Now I deleted product 440799 from the db and loaded it again, now we have 20944 rows and the validator does not fail anymore.
> Would be interesting to know how your data differs from the dump I sent earlier.


### Differences in "raw_flow_yearly" data
The years are different! 

```{r}
# Arrange both dataframes by year, reporter, partner, flow
rawflow0 <- rawflow0 %>% arrange(year, reporter, partner, flow)
rawflow1 <- rawflow1 %>% arrange(year, reporter, partner, flow)
# Check what years are present in both data frames
unique(rawflow0$year)
unique(rawflow1$year)

# Select only common years
rawflow0common <- rawflow0 %>% filter(year %in% 2010:2013) %>% 
    select(year, reporter, partner, flow, quantity, tradevalue)
rawflow1common <- rawflow1 %>% filter(year %in% 2010:2013) %>%
    select(year, reporter, partner, flow, quantity, tradevalue)
# What are the differences now?
common <- rawflow0common %>%
    merge(rawflow1common, 
          by = c("year", "reporter", "partner", "flow"), all = TRUE)
# Variable names in the merged dataset
names(common)
# variables present in one dataset but not in the other
# Not present in rawflow0
sum(is.na(common$tradevalue.x))
# Not present in rawflow1
sum(is.na(common$tradevalue.y))
common <- common %>%
    mutate(partner = iconv(partner, from =  "latin1", to = "UTF-8"),
           notinrawflow0 = is.na(tradevalue.x),
           notinrawflow1 = is.na(tradevalue.y)) %>%
    filter(notinrawflow0 | notinrawflow1)
# 
grepl("ivoire", common$partner)

common %>% kable
```


### Duplicated flows
```{r}
rawflow0 %>% findduplicatedflows %>% nrow
rawflow1 %>% findduplicatedflows %>% 
    select(year, reporter, partner, flow) %>% kable
rawflow1 %>% filter(year == 2013 & reporter == "United Kingdom" &
                partner == "World" & flow == "Re-Import") %>%
    select(year, reporter, partner, flow, quantity, weight, tradevalue, flag) %>%
    kable    
```
